---
documentclass: apa6
classoption: "margin=1in,man,floatsintext"
header-includes:
  # - \usepackage[para,online,flushleft]{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage[singlelinecheck=false]{caption}
  - \captionsetup[subfigure]{singlelinecheck=on, labelfont=normalfont}
  - \captionsetup[figure]{labelfont=it}
  - \shorttitle{SOCIAL MOTIVES \& REINFORCEMENT LEARNING}
title: "Measurement of social motives: relations between self report and reinforcement learning"
author: "John C. Flournoy"
output: 
  pdf_document:
    keep_tex: yes
    template: "default-1.17.0.2_nogeom.tex"
bibliography: "/home/jflournoy/code_new/social-motives-rl-writeup/dissertation.bib"
csl: "/home/jflournoy/Rlibs/probly/bib/apa-old-doi-prefix.csl"
---

```{r setupch4, include=FALSE}
knitr::opts_chunk$set(fig.path = 'figures/', echo = F, warning = F, error = F, message = F)
library(probly)
load('rda/fsmi-cfa.rda')
print_lrt_test_stats <- function(anlrt, result_row = 2, p_sigdig = 3){
    N <- eval(parse(text = paste0('lavaan::nobs(',dimnames(anlrt)[[1]][1], ')')))
    sprintf(paste0('$N = %d$, $\\Delta\\chi^2[',
                   ifelse(round(anlrt[result_row,'Df diff'],3)==1, '%0.0f', '%0.2f'),
                   '] = %0.1f$, $p = %0.', p_sigdig, 'f$'),
            N, anlrt[result_row,'Df diff'], 
            anlrt[result_row,'Chisq diff'], 
            anlrt[result_row,'Pr(>Chisq)'])
}
ggplot2::theme_set(ggplot2::theme_minimal())

sample_abbreviation_note <- 'FCA: foster-care-involved adolescents; CA: community adolescents; CSYA: college students; CSYA-O: college students, online.'
```

# Background

Associations between variables derived from behavior on the Social Probabilistic Learning Task and both self-report trait measures, and self-report motive-relevant behavior were examined in order to test whether individual differences in task response were, as expected, driven by differences in social motives.
For example, if an individual is relatively more driven by mate-seeking concerns (for example, is especially interested in meeting potential romantic partners), the expectation of this paradigm is that this person should also attend relatively more to information in the environment related to the romantic availability of others.
This might result in increased task performance through either decreased random responding (due to overall increased attentiveness) as reflected in the parameter $\xi$, increased sensitivity to each piece of reward feedback ($\rho$), or an increased rate of incorporation of new information ($\epsilon$).
Of course, given that both $\rho$ and $\epsilon$ have non-monotonic relations to performance, a linear relation between mate-seeking motive and either of these parameters could also result in non-monotonic relations to performance.
For this reason, self-report trait measures and motive-relevant (health-related) behaviors will be examined in relation to performance both in the first and last half of the task, as well as with learning model parameters.
There may be relations between overall task ability and social motives as well as other traits, so the minimally social Hungry/Thirsty condition will serve as a baseline by which overall ability on the task is controlled for each individual.

## Self-report measures of social motives

### Fundamental Social Motives Inventory (FSMI)

The FSMI was developed to capture individual differences in social motives that relate to an evolutionary-psychology framework that has as its foundation the idea that motivational systems have adapted to a set of social problems with direct relevance to survival and reproductive fitness [@neel2015].
The inventory contains seven subscales, of which two, the Mate-seeking and Status scales, are relevant to this investigation of the SPLT.
In the paper which introduced this scale, @neel2015 showed that measurements on both of these scales was associated with other motive-relevant self-report instruments, and motive-relevant behaviors.
For example, they found that Mate-seeking was correlated with having asked someone out on a date and Status was correlated with being in a position of leadership at work.
Also, as predicted by the Fundamental Social Motives theoretical framework, @neel2015 found that current relationship status strongly influenced scores on the Mate-seeking questionnaire.
This scale was only administered to the two college samples.

### Kids' Social Reward Questionnaire (K-SRQ)

The original SRQ was developed to measure individual differences in what kinds of social interactions people find rewarding [@foulkes2014].
Understanding what is rewarding and valuable to people is directly related to what people find motivating, with greater reward derived from a particular domain indicating greater motivation in that domain.
In the work that established this measure, validity was investigated using other existing self-report trait scales, with favorable results.
The Sexual Relationships subscale, administered to all samples as a measure mate-seeking motive, correlated strongly with sociosexual orientation measures.
Admiration, administered to all samples as a measure status motive, correlated with dominance, friendliness, hostility, extraversion and narcissism.
Two other subscales, though not focal, are reported in this investigation as well: Passivity, which measures enjoyment of giving control over decisions to others, was negatively correlated with dominance; Sociability, which measures enjoyment of group interactions, showed a pattern of correlations similar to Admiration, but with a correlation to hostility close to zero.
The Kids' SRQ was modified for this study to be more appropriate for adolescents by replacing questions explicitly about sex (e.g., "I enjoy having an active sex life"), with questions that are closely related to sexual and romantic experience (i.e., "I like kissing," "I like having a crush on someone," and "I like flirting").

### Dominance and Presige (D&P) Scale

The Dominance and Prestige scale was developed to measure distinct and non-mutually exclusive strategies for attaining social status [@cheng2013].
Dominance refers to the use of physical or social force to exert control over others such that social rank is increased.
Prestige refers to the process of deploying expertise and performance to garner respect from peers.
These measures were administered to the college sample in part to help reaffirm the validity of the FSMI Status scale, as well as to examine differential associations between SPLT task behavior and particular strategies.
For example, it is plausible that those employing dominance strategies to achieve status need to attend more closely to the social rank, or popularity, of others in order to avoid attempting to exert force upon those more powerful.
For those employing prestige strategies, where social rank is in part a by-product of excellence in other domains, peer popularity may be less important.

### Urgency, Premeditation, Perseverance, and Sensation Seeking (UPPS-P)

The constructs of impulsiveness and sensation seeking figure prominently in research on adolescence [e.g, @collado2014;@harden2011;@mccabe2015;@roberti2004;@romer2010;@shulman2015;@shulman2016;@steinberg2008].
These constructs are also necessarily linked to motivation -- impulsivity and sensation seeking must necessarily select behaviors from those already with some motivational weight attached.
For this reason, the UPPS-P [@whiteside2005], a popular measure of these constructs, was included and used as control variables by which the above measures could be set against as covariates.
This instrument contains scales for both Positive and Negative Urgency, which are supposed to measure impulsivity related to both positive and negative affective states.
It also contains the Sensation Seeking subscale, a measure of enjoyment of activities that pose some amount danger, or uncertainty (i.e., risk).
The Premeditation and Perseverance scales, measuring the tendency, respectively, to think before acting and perseverance in the face of difficulty or tedium, were also included.

## Behavioral outcomes

Examining the relation between these constructs and sexual behavior is done primarily to help establish concurrent validity.
Adolescence is a time when both hormonal and social changes increase propensity to engage in sexual behavior [@udry1988].
These outcomes are developmentally normative, and as such these analyses are not intended as investigations into the relation of mate-seeking motives to adolescent (or early adult) _risk-taking_ behavior.
Of course, these behaviors are health-_relevant_ not just because having sex confers risk of sexually transmitted infection, or pregnancy, but also because successful navigation of romantic and sexual relationships is _developmentally healthy_ [@harden2014a].
In short, sexual and romantic desire, and behavior, should correlate, as well as show developmentally expected increases during the transition from childhood to sexual maturity and adulthood.

Questions about sexual behavior were derived from the the Sexual Experience Survey used by @smith2006, which was itself derived from questions administered by @mcapaldi2002, and include number of recent sexual partners, and sexual encounters.
Risky sexual behavior is assessed by an item asking the participant to report on frequency of a variety of safe sex practices.
In these analyses, this is treated as a health-_risking_ behavior given that these behaviors are the primary means for adults and adolescents alike to ameliorate the risks of developmentally normative sexual behavior.

Alcohol use is another common target outcome for research on adolescent motivation and decisions making, and so provides another touchstone by which we might measure the effectiveness of our instruments, as well as provide additional data on the relations between facets of adolescent motivation and health-relevant behaviors.
Alcohol consumption has been shown to be related to popularity in adolescence [@allen2005], and there is some evidence that alcohol consumption can itself be a strategy for gaining peers' admiration [@mallett2009].
Social gatherings in adulthood and in adolescence are also, more simply, occasions both for navigating social hierarchies, and for drinking alcohol.

As such, in both the college student and full sample, we expect alcohol consumption to be related to measures of motivation to gain status.
We take two outcomes from the Youth Risk Behavior Survey measuring number of days the participant recently had any amount of alcohol, and the number of days where they drank heavily [@centersfordiseasecontrolandprevention2015].
Given that desire for socialization as well as status may motivate attendance at occasions where alcohol is a central feature, associations with the measure of socialization motivation are also examined.
  
# Method

The primary goals of these analyses are to establish that the scales as measures of the constructs of interest, examine whether they provide any convergent evidence that individual differences in the Social Probabilistic Learning Task (SPLT) are related to motives, explore their association with adolescent health-related behavior, and, finally, to explore the associations between this behavior and learning during the SPLT.
Most interpretive decisions will be made on the basis of effect size, without thresholding based on test statistic parametric probabilities.
The primary exception to this is during the psychometric testing phase, as described below.
Analyses of self-report scales will all employ on latent variable models to protect against most types of differential item functioning [@gregorich2006], and to properly account for reliability of measurement [@westfall2016].
All analyses, unless specified otherwise, were performed using the lavaan package [version `r packageVersion('lavaan')`; @rosseel2012] in R [version `r paste0(version$major, '.',version$minor)`; @rcoreteam2018].
Data wrangling and plotting was performed with help of the tidyverse packages [version `r packageVersion('tidyverse')`; @wickham2017], and the psych package [version `r packageVersion('psych')`; @revelle2017].

## Scale validation

Quality of measurement is a crucial precursor to drawing inferences from statistical tests, and so guidelines for testing measurement invariance are followed for each scale below [@gregorich2006].
Invariance between groups based on self-reported gender, sample (when possible), and long-term relationship status (in the college samples) is tested for each self-report scale that is based on a reflective latent variable model (which does not include self-report sexual or substance use behaviors).
Although comparisons between reported gender groups, or samples, are not the focus of this investigation, both groupings appear in models in order to examine the sensitivity of results to group characteristics.

For testing metric invariance, which is required to make inferences based on latent covariance structures, we use the McDonalad Non-centrality Fit Index (MFI) because this has been shown to be robust to differences in sample size and magnitude of item loadings [@kang2016].
In simulations performed by @kang2016, $\Delta\text{MFI} > .01$ was seen in less than 1\% of invariance tests when the null of measurement invariance between groups is true, though the authors caution against rules-of-thumb like this more generally.
In other simulations [@johnflournoy2018], $\Delta\text{MFI} > .012$ was roughly the cut-off for the 1\% level.
The tests below follow these guidelines roughly: when fit differences in the MFI between invariant, or partially-invariant, scales is more than .012, constraints on particular items (suggested by modification indices) are relaxed until the fit difference is no longer substantially concerning.
However, because these items not longer contribute to the model in the same way, and because content coverage of the scales is another important concern for validity [@borsboom2006], the final partially invariant scale is not embraced credulously.
When substantial modification is necessary, sensitivity analyses is performed to compare estimates between the fully constrained scales and the partially constrained (i.e., partially invariant) measurement models.
If results are not too different, then we can be assured that in this case, neither measurement nor content differences are strongly influential on the range of inferences we might make.
If results do appear markedly different, then it is difficult to adjudicate which is the more interpretable.

The final model chosen for more extensive interpretation is selected on the basis of parsimonious model fit.
Smaller values of the Akaike Information Criterion (AIC), an estimate of out-of-sample predictive accuracy that incorporates model parsimony [@akaike1998;@gelman2014], will be used to determine whether emphasis should be given to models that constrain substantive estimates across groups (e.g., samples).
When appropriate, even if the AIC indicates the constrain may provide better predictive accuracy, we can examine differences in estimates between different groups to ensure that this is a reasonable conclusion.

In order to characterize the cross-sectional age associations for each scale, the measurement models will be used to regress each latent variable on a quadratic function of age, with the gender factor interacting with age.
These structural models are intended to provide information on how well these associations converge with expectations based on the social reorientation theory (in the case of status and mate-seeking related constructs), as well as expectations based on past research on impulsivity and sensation seeking during adolescence as described above.
Specifically, the hope, with regard to validating the construct validity of these scales, is to see higher levels of interest in status-related, and mate-seeking-related constructs in older and more pubertally advanced participants (at least up until physical maturity), and higher levels of both sensation seeking and impulse control (or rather, lower levels of urgency) in older participants.
An important caveat to the examination of this descriptive evidence regarding construct validity or more substantive claims is that in cross-sectional data, both convergence and divergence with expectations may be due to factors other than age.

To further examine the case for the validity of these scales as measurements of the constructs they purport to measure, we examine the relations between them.
In order to do so, all latent variable measurement models will be combined to examine inter-scale correlations, with attention to convergence between ostensibly related constructs. 
The college sample was administered all questionnaires, allowing the full inter-scale correlation matrix to be estimated in this sample only.
Correlations among latent variables will also be examined across all samples for the K-SRQ and UPPS-P.
Special attention will be given to the FSMI Mate-seeking and Status scales, the K-SRQ Sexual Relationships and Admiration scales, and the UPPS-P Sensation Seeking scale. 

## Self report relation to SPLT

Four parameters from the reinforcement learning model of SPLT behavior, $\epsilon, \rho, \xi, b$, and two proportions of optimal responses during the first and second half of the run, were correlated with all latent variables in the measurement models established in the step above.
The variables for the SPLT comprise a value for every parameter, for every condition (Hungry/Thirsty, Popular/Unpopular, Dating/Looking), as well as for the contrasts between the two social conditions and the minimally social condition for a total of P = 30.
The contrast variables for model parameters are calculated on the model scale, such that $\epsilon, \xi, \in [0,1]$ and $\rho \in [0,\infty)$, and analyzed without further transformation (differences in proportions are treated similarly).
The parameters from the baseline condition are transformed as they were during model estimation, such that $\epsilon^\prime=\phi^{-1}(\epsilon)$, $\xi^\prime=\phi^{-1}(\xi)$, $\rho^\prime = ln(\rho)$, where $\phi^{-1}$ is the quantile function for the unit normal density.
The condition-specific (but not contrast) values for proportion of optimal responses were also transformed using the quantile function.

Only the contrasts between the social motive and baseline conditions, and the baseline condition itself are of interest.
In particular, the correlations between the SPLT contrast variables and self-report measures are expected to indicate the degree to which motive effects on learning are related to individual differences in self-report motives.
The relation between SPLT variables in the baseline condition are relevant for interpreting the possible explanations for individual differences in reinforcement learning more generally.
Interpretation of the results will focus on the size and sign of the correlation with regard to the precision as suggested by the 95\% confidence intervals.


## Self report relation to outcomes

First, the motive-related scales may be further evaluated for construct validity on the basis of their relation to self-reported behaviors which would be expected to result from those motives.
For example, in general, those who are in long-term relationships should report lower levels of interest in seeking out new romantic or sexual partners (i.e., lower scores on the FSMI Mate-Seeking scale).
Among all participants, higher levels on the K-SRQ Sexual Relationships scale should be related to higher numbers of sexual encounters (specifically, "how many times, in the last six months, have you had sex/sexual relations," with sexual relations defines as "oral, vaginal or anal, and can be with male or female partners").
Among those who are not in long-term relationships, higher FSMI Mate-Seeking and K-SRQ Sexual Relationships (which is supposed to measure the enjoyment of sexual behavior) should be related to having had more, recent, sexual partners (specifically, "with how many different people in the last 6 months").
Finally, though it is more speculative, because of the item content related to parties and social events, higher scores on the K-SRQ Sociability may be expected to relate to greater number of days having had at least one drink ("In the past 30 days, on how many days did you have at least one drink of alcohol"), at least in the college sample.

The more substantively interesting questions involve the relation between risky sexual behavior and alcohol use.
The outcomes of interest are:

- "In the last 6 months, what percentage of the time when you've had sexual relations (oral, vaginal or anal) did you USE safe sex practices to prevent sexually transmitted diseases - condoms, blood tests, dental dams, etc.)"
- "In the past 30 days, on how many days did you have 5 or more drinks of alcohol in a row, that is, within a couple hours?"

The data generating processes for the behavioral outcomes are not Gaussian, nor do they necessarily result in data where a unit difference between responses has the same meaning.
To account for this, certain analysis and data transformation decisions were made.
For frequency of sex, number of partners, number of past 30 days with alcohol, and number of past 30 days with five or more alcoholic drinks, responses were analyzed as ordinal outcomes.
This required binning responses for some variables.
Participants were free to input any number for the frequency of sex question, so to ensure that ordered categories contained at least 30 participants, responses were binned into groups with 0, 1-9, 10-19, 20-39, and 40 or more sexual encounters in the past six months.
Responses to the number of partners in the last six months was also binned into groups with 0, 1, 2, and 3 or more, with at least 40 participants in each bin.
The response options for both alcohol use questions were constrained, with response options, in days, set to 0, 1-2, 3-5, 6-9, 10-19, 20-29, and "All 30".
For both responses, very few or no participants selected 20 or more days, and only 6 participants selected 10-19 days for the question asking about having had five or more drinks in one session.
Bins on the high end were combined resulting in five bins for the question asking about days having any number of drinks (0, 1-2, 3-5, 6-9, 10 or more), and in four bins for the question about days having five or more drinks (0, 1-2, 3-5, 6 or more).
The question regarding safe sex practice was analyzed continuously because the response options asked about continuous frequency, and the question about substance use prior to sex was analyzed as a binary categorical outcome.
Analysis of categorical outcomes used ordinal regression using a diagonally weighted least squares estimator with robust standard errors and a mean and variance adjusted test statistic (corresponding to the WLSMV estimator in the lavaan package).
This estimation method requires listwise deletion for any cases without complete responses on the questionnaires.
This results in lower samples sizes for analyses comparing the UPPS-P scales as described below.

Gender and (if the adolescent participants could be included) age were added as covariates to all models.
Associations between target motive variables and outcomes were estimated both with and without UPPS-P Sensation Seeking and both Urgency variables as covariates to explore sensitivity of the outcomes to variance shared between these measurements.
Also, in the case of K-SRQ variables, sensitivity to the inclusion of highly correlated subscales was also tested.
For example, if the Sexual Relationships variable was included as the focal independent variable, the Admiration and Sociability variables were added as covariates.

Interpretation of associations (both with self-report scales, and when considering the SPLT variables) with self report behavior is based on the sign of the focal motive variable coefficient, and on the $\chi^2$ difference test ($\alpha = .005$) between a model including the focal motive variable and covariates, and the same model with the motive variable coefficient fixed to 0.
To examine sensitivity to inclusion of the UPPS-P scales, comparison was between the model with all variables (motive, UPPS-P Sensation Seeking, and Positive and Negative Urgency), and the same model with the motive coefficient fixed to 0.
That is, the test examined fit decrease due to excluding the relation between the outcome and motive variable.
Sensitivity to the addition of K-SRQ subscales tested models with the zero-coefficient constraint set on the added variable, with examination of size and sign change of the focal motive variable coefficient.

Finally, the subset of these analyses examining associations with number of sexual partners are estimated using only the participants who report not being in a long-term relationship.
For completeness, the same analysis is repeated without regard for long-term relationship status and note is made of any substantial differences.

## SPLT relation to outcomes

The same behavioral outcomes are also regressed on the relevant contrasts of all SPLT variables.
Specifically, self report sexual behavior is regressed on the Dating/Looking - Hungry/Thirsty contrast for each parameter, $\epsilon, \rho, \xi, b$, and the same contrast for the proportion of optimal decisions made during both the first and last half of the run.
Self-report drinking is regressed on the Popular/Unpopular - Hungry/Thirsty contrasts of these variables.
An initial model is fit to a limited sample that includes just unpartnered college students (sexual behavior outcomes), or just college students (alcohol consumption outcomes), and another model is fit using both college and both adolescent samples.
Gender and age (when the both college and adolescent samples are combined) are included as covariates.
The test of association between the SPLT variable and the outcome is, as above, an adjusted $\Delta\chi^2$ test, with decision cut-off of $\alpha = .005$ (note, however, that this procedure was not established _a priori_).

# Results

## Evaluating scales

Descriptive statistics including mean, standard deviation, Cronbach's $\alpha$, and Guttman's $\lambda\text{-6}$ coefficient are provided for each computed scale, for each sample.
These statistics were computed as means of all items responded to, including cases with missing values, and with no imputation, using the alpha function in the psych package [@revelle2017].
Though all scales will be evaluated using latent variable measurement models, but these statistics may provide a helpful starting point of reference.
Statistics for scales administered to the college samples, only, appear in Table \ref{tab:srdesccollege}, and for the K-SRQ and the UPPS-P, administered to the adolescent samples only, can be found in Tables \ref{tab:srdescksrq} and \ref{tab:srdescuppsp}.

\begin{table}
\centering
\begin{threeparttable}
\caption{Descriptive statistics for self-report scales, college-samples} \label{tab:srdesccollege}
\footnotesize
```{r , results='asis'}
knitr::kable(scale_descriptives_table_col,
             col.names = c('Scale name', 'Sample', '$\\alpha$', '$\\lambda\\text{-6}$', 'Mean', 'SD'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '',
             digits = 2)
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item SD = Standard Deviation;F = Fundamental Social Motives Inventory; D = Dominance and Prestige Scale.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}
\centering
\begin{threeparttable}
\caption{Descriptive statistics for K-SRQ} \label{tab:srdescksrq}
\footnotesize
```{r , results='asis'}
knitr::kable(scale_descriptives_table_comb_ksrq,
             col.names = c('Scale name', 'Sample', '$\\alpha$', '$\\lambda\\text{-6}$', 'Mean', 'SD'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '',
             digits = 2)
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item SD = Standard Deviation; K = Kids' Social Reward Questionnaire.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}
\centering
\begin{threeparttable}
\caption{Descriptive statistics for UPPS-P} \label{tab:srdescuppsp}
\footnotesize
```{r , results='asis'}
knitr::kable(scale_descriptives_table_comb_uppsp,
             col.names = c('Scale name', 'Sample', '$\\alpha$', '$\\lambda\\text{-6}$', 'Mean', 'SD'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '',
             digits = 2)
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item SD = Standard Deviation; U = Urgency, Premeditation, Perseverance, Sensation Seeking Scale.
\end{tablenotes}
\end{threeparttable}
\end{table}




### Fundamental Social Motives Inventory

The Fundamental Social Motives Inventory (FSMI) was administered to the college sample only (both because the questions are not appropriate for younger adolescents, and because time did not allow for this additional inventory).
For the specific purposes of this study, the analyses below focus on the Mate-seeking and Status factors, though an examination of the full inventory is doubtless of interest, more generally.
Invariance of loadings of items on factors (metric invariance) between participants reporting male versus female gender on the pubertal development scale (PDS) was tested and found acceptable, with a slight increase in fit after constraining factor loadings ($\Delta\text{MFI}= `r round(fsmi_matestat_cfa_fit[['delta1']][['mfi']],3)`$).
The final confirmatory factor analysis (CFA) for mate-seeking and status factors fit well ($\text{N}_{\text{total}}$ = `r lavaan::inspect(fsmi_cfa_matestat_gender_metric, what = 'ntotal')`, RMSEA = `r round(fsmi_matestat_cfa_fit[['fsmi_cfa_matestat_gender_metric']][['rmsea']],3)`, $\text{MFI}$ = `r sprintf('%0.2f', fsmi_matestat_cfa_fit[['fsmi_cfa_matestat_gender_metric']][['mfi']])`, $\hat\gamma$ = `r round(fsmi_matestat_cfa_fit[['fsmi_cfa_matestat_gender_metric']][['gammaHat']],2)`).
Adding equality constraints on the latent covariance structure resulted in better parsimony-adjusted model fit as measured by the differences in the Akaike information criterion (AIC; $\Delta\text{AIC}= `r round(fsmi_matestat_cfa_fit[['delta2']][['AIC']],0)`$), with the two factors showing a very small, positive, non-significant correlation ($r = `r round(fsmi_mate_stat_cor[['std.all']][[1]],2)`, Z = `r round(fsmi_mate_stat_cor[['z']][[1]],2)`$).
Note that item 54, "I do not worry very much about losing status," has a particularly low loading on the status factor ($\beta$ = `r round(fsmi_mate_stat_par[fsmi_mate_stat_par$op == '=~' & fsmi_mate_stat_par$rhs == 'fsmi_qs_54','std.all'][[1]],2)`).
Associations with age and pubertal development were not examined given the narrow range of each of these measurements in this sample.

### Dominance and Prestige

The Dominance and Prestige scale (D&P) was also administered to the college sample only (because time did not allow for it).
Invariance of loadings between males and females was tested and found to decrease fit ($\Delta\text{MFI} = `r round(dnp_cfa_fit[['delta1']][['mfi']],3)`$).
Modification indices indicated that freeing the constraint of equality between groups for the item, "Some people are afraid of me", would improve fit, and doing so resulted in an acceptable decrease in fit between the (partially) metric invariant and unconstrained models ($\Delta\text{MFI} = `r round(dnp_cfa_fit[['delta2']][['mfi']],3)`$).
Further improvements were not sought as relaxing constraints on this single item both broadly protects the scale's content coverage [@borsboom2006], while providing fairly acceptable fit differences. 
This partial-metric invariance model is the model used in all analyses below, and sensitivity of results to the relaxation of this constraint is not tested.
The final, partially metric invariant CFA for Dominance and Prestige factors fit somewhat poorly ($\text{N}_{\text{total}}$ = `r lavaan::inspect(dnp_cfa_gender_partmetric, what = 'ntotal')`, RMSEA = `r round(dnp_cfa_fit[['dnp_cfa_gender_partmetric']][['rmsea']],3)`, $\text{MFI}$ = `r round(dnp_cfa_fit[['dnp_cfa_gender_partmetric']][['mfi']],2)`, $\hat\gamma$ = `r round(dnp_cfa_fit[['dnp_cfa_gender_partmetric']][['gammaHat']],3)`).
Further constraining the latent covariance and variance to be equal between both females and males resulted in slightly better model fit ($\Delta\text{AIC} = `r round(dnp_cfa_fit[['delta3']][['AIC']],0)`$), with the two factors showing a small, positive, correlation ($r = `r round(dnp_mate_stat_cor[['std.all']][[1]],2)`, Z = `r round(dnp_mate_stat_cor[['z']][[1]],2)`$).
Associations with age and pubertal development were not examined given the narrow range of each of these measurements in this sample.

### Kids' Social Reward Questionnaire

The Kids' Social Reward Questionnaire (K-SRQ), and remaining self-report instruments, were administered to each college and adolescent samples.
The Negative Social Potency and Prosocial Interactions factor indicators were almost all very highly skewed, so these subscales were dropped before any further analysis was performed.
The four factors included below are: Admiration, Passivity, Sexual Relationships, and Sociability.

A CFA for the four K-SRQ factors was tested for metric invariance across samples by comparing a model with loadings constrained to equality to one with loadings free to vary across the four groups.
This resulted in considerably worse model fit ($\Delta\text{MFI} = `r round(ksrq_cfa_fit[['delta1a']][['mfi']],3)`$).
Loading constraints were removed from four items ("I like it if others looks up to me", "I like being a member of a group/club", "I like kissing", "I like flirting") that load onto Admiration, Sociability, and  Sexual Relationships, respectively.
This resulted in adequately small fit differences between the unconstrained and partial-metric invariance models (`r ksrq_cfa_metric_invar_stats`).
The final model showed somewhat poor fit ($\text{N}_{\text{total}}$ = `r lavaan::inspect(ksrq_cfa, what = 'ntotal')`, RMSEA = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['rmsea']],2)`, MFI = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['mfi']],2)`, $\hat\gamma$ = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['gammaHat']],3)`). 

Metric invariance was also evaluated between reported gender groups.
The partial-metric invariance model from above was compared to the model without loadings constrained.
The fit decrease between these two models was not too big ($\Delta\text{MFI} = `r round(ksrq_cfa_fit[['delta_gender']][['mfi']],3)`$).

Next, models with and without between-sample equality constraints on latent variances and covariances were compared, resulting in better fit ($\Delta\text{AIC} = `r round(ksrq_cfa_fit$delta2[['AIC']])`$) for the constrained model.
Loadings without constraints across groups (in the partial-metric invariance model) do not contribute to this comparison, so the magnitude of correlations was compared across the partial-metric invariant and fully constrained models (both with the constraint of covariance across groups).
The differences in correlations were in the range _r_ = [`r round(min(ksrq_cor_diff),2)`, `r round(max(ksrq_cor_diff),2)`], indicating very low sensitivity the constraints on item loadings.
To maintain content coverage, the fully constrained model, with equality of loadings across all samples, is used, though sensitivity to constraints on the four non-invariant loadings is assessed again in subsequent analyses.

Correlations among the Admiration, Sociability, and  Sexual Relationships subscales were all high ($`r sprintf('r = [%.2f, %.2f]', min(lavaan::inspect(ksrq_cfa_metric_cov, 'cor.lv')$yads_online[-2, -2]), max(lavaan::inspect(ksrq_cfa_metric_cov, 'cor.lv')$yads_online[-2, -2] - diag(3)))`$). 
Correlations of Passivity with these three subscales were generally small, but positive ($`r sprintf('r = [%.2f, %.2f]', min((lavaan::inspect(ksrq_cfa_metric_cov, 'cor.lv')$yads_online)[2,]), max((lavaan::inspect(ksrq_cfa_metric_cov, 'cor.lv')$yads_online - diag(4))[2,]))`$; see Tables \ref{tab:collegecor} and \ref{tab:allcor} for all latent variable correlations).
This pattern of correlation is similar to what was seen in the scale construction sample [@foulkes2014].

#### Age

Variation in latent factors was described by a quadratic age regression, with age effects allowed to vary by gender.
The partial metric invariance model was used to test whether constraining regression coefficients across samples resulted in a better fitting model as assessed by the AIC.
Constraining the regression coefficients across sample resulted in reduced AIC ($\Delta\text{AIC} = `r round(diff(ksrq_cfa_metric_agesem_fit[,'AIC']),0)`$), indicated better model fit.
Standardized coefficients were extremely similar between the model with partial metric invariance and the model with all loadings constrained to equality across samples ($\Delta\beta = `r sprintf('[%0.2f, %0.2f]', min(ksrq_agesem_betadiffs), max(ksrq_agesem_betadiffs))`$), so the fully constrained model is interpreted.
Model-expected trend lines show somewhat higher scores in late adolescent participants (Figure \ref{fig:ksrqagepredplot}), with the largest quadratic effects (averaged across gender) for Admiration, Sociability, and Sexual Relationships (Table \ref{tab:ksrqage}).
Gender differences in the linear model coefficients were fairly small.

```{r ksrqagepredplot, fig.width=5.875, fig.height=3, include=F}
ksrq_age_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/ksrqagepredplot-1.pdf}
\caption[K-SRQ model-predicted PDS associations]{K-SRQ model-predicted PDS associations. Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:ksrqagepredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{K-SRQ factors regressed on age and gender} \label{tab:ksrqage}
\footnotesize
```{r ksrqage, results='asis'}
knitr::kable(dplyr::mutate(ksrq_std_par_table, 
                           lhs = sub('Sexual Relationships', 'Sexual rel.', lhs)),
             col.names = c('K-SRQ factor', 'Sample', 
                           '$\\text{Age}$', '$\\text{Age}\\times\\text{Gender}$', 
                           '$\\text{Age}^2$', '$\\text{Age}^2\\times\\text{Gender}$'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '')
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95\% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}

#### Puberty

We also described variation in latent factors by a quadratic PDS regression, with effects allowed to vary by gender.
The partial metric invariance model with constrained regression coefficients across samples resulted in a better fit ($\Delta\text{AIC} = `r round(diff(ksrq_cfa_metric_pdssem_fit[,'AIC']),0)`$).
Again, standardized coefficients were extremely similar between the model with partial metric invariance and the model with all loadings constrained to equality across samples ($\Delta\beta = `r sprintf('[%0.2f, %0.2f]', min(ksrq_pdssem_betadiffs), max(ksrq_pdssem_betadiffs))`$), so the fully constrained model is interpreted.
Model-expected trend lines show somewhat lower scores associated with lower PDS, primarily for female participants (Figure \ref{fig:ksrqpdspredplot}), with the largest linear effects (averaged across gender) for Sociability, and Sexual Relationships (Table \ref{tab:ksrqpds}).
Though a model including the quadratic effect of PDS was estimated, this was done to allow deceleration in the linear trend to approximate the observed asymptotic relation.

```{r ksrqpdspredplot, fig.width=5.875, fig.height=3, include=F}
ksrq_pds_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/ksrqpdspredplot-1.pdf}
\caption[K-SRQ model-predicted PDS associations]{K-SRQ model-predicted age associations. Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:ksrqpdspredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{K-SRQ factors regressed on age and gender} \label{tab:ksrqpds}
\footnotesize
```{r ksrqpds, results='asis'}
knitr::kable(dplyr::mutate(ksrq_std_pds_par_table, 
                           lhs = sub('Sexual Relationships', 'Sexual rel.', lhs)),
             col.names = c('K-SRQ factor', 'Sample', 
                           '$\\text{PDS}$', '$\\text{PDS}\\times\\text{Gender}$', 
                           '$\\text{PDS}^2$', '$\\text{PDS}^2\\times\\text{Gender}$'),
             booktabs = TRUE,
             escape = F, format = 'latex',
             linesep = '')
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95\% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}
  
### Urgency, Premedidation, Perseverence, Sensation Seeking and Positive Urgency

The Urgency, Premeditation, Perseverance, Sensation Seeking and Positive Urgency (UPPS-P) scale was also administered to all four samples.
This instrument uses a large number of indicators for each latent factor.
When doing an analysis with a separate group for each sample, this results in a number of variables that greater than the number of observations.
For this reason, measurement is assessed between combined adolescent and combined college-student samples.

A CFA for the five UPPS-P factors was tested for metric invariance across age groups by comparing a model with loadings constrained to equality to one with loadings free to vary across the four groups.
This resulted in adequately small fit differences between the unconstrained and invariant models (`r uppsp_cfa_metric_invar_stats`).
Metric invariance was also evaluated between reported gender groups, with practically no fit decrease ($`r sprintf('\\Delta\\text{MFI} = %0.4f', round(uppsp_cfa_fit[['delta_gender']][['mfi']],4))`$).

The fit statistics show mixed results, with acceptable RMSEA but also low MFI and $\hat\gamma$ ($\text{N}_{\text{total}}$ = `r lavaan::inspect(uppsp_cfa_metric, what = 'ntotal')`, RMSEA = `r round(uppsp_cfa_fit$uppsp_cfa_metric[['rmsea']],2)`, MFI = `r round(uppsp_cfa_fit$uppsp_cfa_metric[['mfi']],2)`, $\hat\gamma$ = `r round(uppsp_cfa_fit$uppsp_cfa_metric[['gammaHat']],3)`).
Other work has reported similarly ambiguous fit statistics for the UPPS-P in confirmatory factor analysis [@cyders2013].
This may be the reason for relatively similar degrees of fit between the unconstrained and metric invariance models.
We proceed with these analyses with the caveat that we cannot be confident latent variable covariance is not contaminated by differential item use.

Models with and without constraints on the equality of latent variances and covariances between age-groups were compared, resulting in lower AIC ($\Delta\text{AIC} = `r round(uppsp_cfa_fit$delta2[['AIC']])`$).
Absolute magnitude of correlations among all scales were between small and large (_r_ = `r sprintf('[%0.2f, %0.2f]', min(abs(uppsp_factor_cor)[abs(uppsp_factor_cor)>0]), max(abs(uppsp_factor_cor)))`), with very a very large correlation between Negative and Positive Urgency (_r_ = `r round(pos_neg_urg,2)`; see Tables \ref{tab:collegecor} and \ref{tab:allcor} for all latent variable correlations). 
Scales keyed in the positively valenced direction (Premeditation and Perseverance) were negatively correlated with negatively valenced scales (Negative and Positive Urgency, and Sensation Seeking).

#### Age

As with the K-SRQ, the association between UPPS-P factors and age was examined by regressing the latent variables on the interaction between a quadratic age model and gender.
Again, constraining the regression coefficients across sample results in reduced AIC ($\Delta\text{AIC} = `r round(diff(uppsp_cfa_metric_agesem_fit[,'AIC']),0)`$), indicating better model fit.
The model expectations reveal that the association between age and factor score may be conditional on reported gender.
Plots for Negative and Positive Urgency, Perseverance, and Sensation seeking show diverging quadratic trends for females and males (though note that expected correlation with age is fairly flat for both genders for Premeditation; Figure \ref{fig:uppspagepredplot}).
The standardized effects indicate that, averaged across gender, linear and quadratic associations were very small, but that for Perseverance, and Sensation seeking, the gender difference in the quadratic term is especially robust (Table \ref{tab:uppspage}).

```{r uppspagepredplot, fig.width=5.875, fig.height=3, include=F}
uppsp_age_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/uppspagepredplot-1.pdf}
\caption[UPPS-P model-predicted age associations]{UPPS-P model-predicted age associations. Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:uppspagepredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{UPPS-P factors regressed on age and gender} \label{tab:uppspage}
\footnotesize
```{r uppspage, results='asis'}
knitr::kable(uppsp_std_par_table,
             col.names = c('UPPS-P factor', 'Sample', 
                           '$\\text{Age}$', '$\\text{Age}\\times\\text{Gender}$', 
                           '$\\text{Age}^2$', '$\\text{Age}^2\\times\\text{Gender}$'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '')
```
\begin{tablenotes}
\footnotesize
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95\% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}

#### Puberty

The association between the UPPS-P latent factors and PDS was described using the same procedures as above.
Constraining the regression coefficients across sample results in better fit ($\Delta\text{AIC} = `r round(diff(uppsp_cfa_metric_pdssem_fit[,'AIC']),0)`$).
Model expected trend lines show very flat relations between PDS and latent factor scores with the exception of an accelerating quadratic fit and negative linear slope (tangent at PDS = 3) for Positive Urgency (Figure \ref{fig:uppsppdspredplot}).
Standardized regression coefficients echo the plots, all having small magnitude, with the exception of the larger negative slope for Positive Urgency (Table \ref{tab:uppsppds}).

```{r uppsppdspredplot, fig.width=5.875, fig.height=3, include=F}
uppsp_pds_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/uppsppdspredplot-1.pdf}
\caption[UPPS-P model-predicted PDS associations]{UPPS-P model-predicted PDS associations. Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:uppsppdspredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{K-SRQ factors regressed on age and gender} \label{tab:uppsppds}
\footnotesize
```{r uppsppds, results='asis'}
knitr::kable(uppsp_std_pds_par_table,
             col.names = c('UPPS-P factor', 'Sample', 
                           '$\\text{PDS}$', '$\\text{PDS}\\times\\text{Gender}$', 
                           '$\\text{PDS}^2$', '$\\text{PDS}^2\\times\\text{Gender}$'),
             booktabs = TRUE, format = 'latex', escape = F,
             linesep = '')
```
\begin{tablenotes}
\footnotesize
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95\% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}

### All scale correlations

All measurement models were combined in order to estimate correlations among the latent factors.
Though each of the measurement models show acceptable fit, this combine model shows very poor fit (college samples: `r sprintf('$N = %d, \\text{RMSEA} = %0.2f, \\text{MFI} = %0.0f$', lavaan::inspect(all_scales_college, 'ntotal'), lavaan::fitmeasures(all_scales_college, 'rmsea')[[1]], lavaan::fitmeasures(all_scales_college, 'mfi')[[1]])`; combined samples: `r sprintf('$N = %d, \\text{RMSEA} = %0.2f, \\text{MFI} = %0.0f$', lavaan::inspect(all_scales_adol, 'ntotal'), lavaan::fitmeasures(all_scales_adol, 'rmsea')[[1]], lavaan::fitmeasures(all_scales_adol, 'mfi')[[1]])`), indicating that this set of latent variables does not reproduce well the observed covariance among the indicators.
The possible reasons for this are discussed later, below, but not further explored analytically.
Combine models are described within the full college sample for all questionnaires, with a combined model with the K-SRQ and the UPPS-P across all four samples.

#### College samples

As was done above, the model constrained all loadings (except those freed above) and the latent covariance structure to be equal across gender groups, which fit better than allowing covariance to vary between genders ($\Delta\text{AIC} = `r round(diff(AIC(all_scales_college, all_scales_college_cor)[,2]))`$).
Further constraining the model to have equal loadings for all items resulted in only trivial differences in the latent variable correlation magnitudes (maximum absolute $\Delta r = `r round(college_cor_sensitivity_gender_abs_max, 2)`$).
This same pattern held when examining differences across partnered and unpartnered status ($\Delta\text{AIC} = `r round(diff(AIC(all_scales_college_partnered, all_scales_college_partnered_cor)[,2]))`$; maximum absolute $\Delta r = `r round(college_cor_sensitivity_partnered_abs_max, 2)`$).

The primary variables of interest are FMSI Mate-seeking, FSMI Status, K-SRQ Admiration, K-SRQ Sexual Relationships, and UPPS-P Sensation seeking.
These scales were, _a priori_, intended to measure mate-seeking and status motivation.
All latent variable correlations for the comprehensive measurement battery administered to the college samples can be found in Table \ref{tab:collegecor}.
The FSMI Mate-seeking variable shows correlations in the small-to-moderate range with most of the other variables, including both K-SQR Sexual Relationships and UPPS-P Sensation Seeking.
The FSMI Status variable showed more robust correlations with closely related variables.
Correlations with Dominance and Prestige status-strategies were large and very large, respectively, while the correlation with admiration was also very large.
Interestingly, the association between FSMI Status and K-SRQ Sexual Relationships was somewhat large.
The correlation between FSMI Status and K-SRQ Sociability is also large.
The strength of association between K-SRQ Admiration and Sexual Relationships is extremely large, as is the correlation between these variables and the K-SRQ Sociability scale.
Sensation Seeking shows medium correlations with the FMSI Status variable, both Dominance and Prestige variables, as well as K-SRQ Sexual Relationships and Sociability scales, although the size of these correlations is of roughly the same magnitude as the correlations between Sensation Seeking and other UPPS-P variables.


\begin{table}
\centering
\begin{threeparttable}
\caption{Latent variable correlations, college sample} \label{tab:collegecor}
\footnotesize
\centering
```{r collegecor, results='asis'}
latent_var_labels <- c(
  ' 1. fsmi mate' = ' 1. Mate-Seeking\\textsuperscript{F}',
  ' 2. fsmi stat' = ' 2. Status\\textsuperscript{F}',
  ' 3. dominance score' = ' 3. Dominance\\textsuperscript{D}',
  ' 4. prestige score' = ' 4. Prestige\\textsuperscript{D}',
  ' 5. neg urgency' = ' 5. Negative Urgency\\textsuperscript{U}',
  ' 6. premeditation' = ' 6. Premeditation\\textsuperscript{U}',
  ' 7. perseverance' = ' 7. Perseverance\\textsuperscript{U}',
  ' 8. sensation seeking' = ' 8. Sensation Seeking\\textsuperscript{U}',
  ' 9. pos urgency' = ' 9. Positive Urgency\\textsuperscript{U}',
  '10. k srq admiration' = '10. Admiration\\textsuperscript{K}',
  '11. k srq passivity' = '11. Passivity\\textsuperscript{K}',
  '12. k srq sexual relationships' = '12. Sexual Relationships\\textsuperscript{K}',
  '13. k srq sociability' = '13. Sociability\\textsuperscript{K}')
latent_var_cor_table <- dplyr::mutate_at(latent_var_cor_table,
                                         dplyr::vars(2:14),
                                         dplyr::funs(ifelse(is.na(.), '', .)))
knitr::kable(
  dplyr::mutate(latent_var_cor_table[c(1:8)],
                rhs = latent_var_labels[rhs]),
  booktabs = TRUE, format = 'latex', escape = F,
  col.names = c('Latent variable', 1:7),
  linesep = '',
  align = c('l', rep('r', 7)))
```
\centering
```{r collegecor2, results='asis'}
knitr::kable(
  dplyr::mutate(latent_var_cor_table[c(1,9:14)],
                rhs = latent_var_labels[rhs]),
  booktabs = TRUE, format = 'latex', escape = F,
  col.names = c('Latent variable', 8:13),
  linesep = '',
  align = c('l', rep('r', 6)))
```
\begin{tablenotes}
\footnotesize
\item F = Fundamental Social Motives Inventory; D = Dominance \& Prestige scale; U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire.
\item Correlations between latent variables appear in the lower triangle; 95\% confidence intervals appear in the upper triangle. Source model constrains correlations and loadings for all items to be equal across gender groups.
\end{tablenotes}
\end{threeparttable}
\end{table}

#### College and adolescent samples

As was done above, the model constrained all loadings (except those freed above) and the latent covariance structure to be equal across subsamples, which fit better than allowing covariance to vary between them ($\Delta\text{AIC} = `r round(diff(AIC(all_scales_adol, all_scales_adol_cor)[,2]))`$).
Further constraining the model to have equal loadings for all items resulted in only trivial differences in the latent variable correlation magnitudes (maximum absolute $\Delta r = `r round(adol_cor_sensitivity_abs_max, 2)`$).

Of the variables measured in all four samples, those of primary interest are K-SRQ Admiration, K-SRQ Sexual Relationships, and UPPS-P Sensation seeking.
The first two scales were, _a priori_, intended to measure mate-seeking and status motivation, while the UPPS-P Sensation Seeking scale is intended as a reference point for other research on adolescence.
All correlations across the full sample can be found in Table \ref{tab:allcor}.
Across all samples, as with just the college sample, the correlations among the K-SRQ Admiration, Sexual Relationships, and Sociability scales are extremely large.
Sensation Seeking shows slightly stronger correlations with K-SRQ Admiration, Sexual Relationships and Sociability scales than in the college sample alone. 
However, as with the college sample, above, the size of these correlations are similar to, or smaller than, the UPPS-P scale inter-questionnaire correlations.

Even though model fit was improved by constraining correlations to be the same across samples, examining the pattern of estimated within-sample correlations may help increase our confidence that this is reasonable.
It is fairly clear in Figure \ref{fig:allscalesadolplot} that, with some variability, the same high level of correlation among three of the four K-SRQ variables is present in all samples, though that between Admiration and Sexual Relationships may be somewhat attenuated for the in-lab college students (CSYA) and the community adolescents (CA).
The relation between UPPS-P Sensation Seeking and K-SRQ Admiration, Sexual Relationships, and Sociability seems quite consistent.
The most striking differences are tangential to the scope of this manuscript, but may be worth pointing out: the correlation between KSR-Q Passivity and both Sexual Relationships and Admiration is large and negative within the foster-care-involved adolescent sample (FCA), whereas for the other samples it is usually closer to zero (with community adolescents [CA] showing a large positive correlation between Passivity and Sexual Relationships).

\begin{table}
\centering
\begin{threeparttable}
\caption{Latent variable correlations, all samples} \label{tab:allcor}
\footnotesize
\centering
```{r allcor, results='asis'}
latent_var_labels_all <- c(
  ' 1. neg urgency' = ' 1. Negative Urgency\\textsuperscript{U}',
  ' 2. premeditation' = ' 2. Premeditation\\textsuperscript{U}',
  ' 3. perseverance' = ' 3. Perseverance\\textsuperscript{U}',
  ' 4. sensation seeking' = ' 4. Sensation Seeking\\textsuperscript{U}',
  ' 5. pos urgency' = ' 5. Positive Urgency\\textsuperscript{U}',
  ' 6. k srq admiration' = ' 6. Admiration\\textsuperscript{K}',
  ' 7. k srq passivity' = ' 7. Passivity\\textsuperscript{K}',
  ' 8. k srq sexual relationships' = ' 8. Sexual Relationships\\textsuperscript{K}',
  ' 9. k srq sociability' = ' 9. Sociability\\textsuperscript{K}')
latent_var_cor_table_adol <- dplyr::mutate_at(latent_var_cor_table_adol,
                                              dplyr::vars(2:10),
                                              dplyr::funs(ifelse(is.na(.), '', .)))
knitr::kable(
  dplyr::mutate(latent_var_cor_table_adol[c(1:6)],
                rhs = latent_var_labels_all[rhs]),
  booktabs = TRUE, format = 'latex', escape = F,
  col.names = c('Latent variable', 1:5),
  linesep = '',
  align = c('l', rep('r', 5)))
```
\centering
```{r allcor2, results='asis'}
knitr::kable(
  dplyr::mutate(latent_var_cor_table_adol[c(1,7:10)],
                rhs = latent_var_labels_all[rhs]),
  booktabs = TRUE, format = 'latex', escape = F,
  col.names = c('Latent variable', 6:9),
  linesep = '',
  align = c('l', rep('r', 4)))
```
\begin{tablenotes}
\footnotesize
\item U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire.
\item Correlations between latent variables appear in the lower triangle; 95\% confidence intervals appear in the upper triangle. Source model constrains correlations and loadings for all items to be equal across gender groups.
\end{tablenotes}
\end{threeparttable}
\end{table}


```{r allscalesadolplot, fig.width=5.875, fig.height=7, include=F}
all_scales_adol_plot + 
  ggplot2::theme(strip.text.y = ggplot2::element_text(size = 7))
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/allscalesadolplot-1.pdf}
\caption[Latent variable correlations by subsample]{Latent variable correlations by subsample. The correlation for all subsamples is taken from the model with all loadings and latent (co)variances constrained to be equal across subsamples. Subsample correlations are taken from the model with partial metric invariance. U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire; - = Negative; + = Positive. `r sample_abbreviation_note`}
\label{fig:allscalesadolplot}
\end{figure}

## Self report relation to SPLT

Four parameters from the reinforcement learning model of the SPLT behavior, and two summary statistics of optimal responding during the first and second half of the run, were correlated with all latent variables.
The variables for the SPLT have a value for every condition (Hungry/Thirsty, Popular/Unpopular, Dating/Looking), as well as for the contrasts between the two social conditions and the minimally social condition.
Including all SPLT variables in the combined measurement model caused problems with convergence that were not solved by changing the optimizer or adjusting optimization control options.
For this reason, latent variable correlations were estimated for each SPLT-variable parameter subset separately.

### College sample

```{r}
scale_splt_cor_df_collegeminmax <- dplyr::summarize(
  dplyr::filter(
    scale_splt_cor_df_college,
    grepl('HngT_fac$', rhs),
    !grepl('^b_', rhs)), 
  min=min(est.std), max=max(est.std))

scale_splt_cor_df_collegeminmaxbl <- dplyr::summarize(
  dplyr::filter(
    scale_splt_cor_df_college,
    grepl('HngryThrsty_fac$', rhs),
    !grepl('^b_', rhs)), 
  min=min(est.std), max=max(est.std))
```

Correlations were first examined in the college sample (N = `r lavaan::nobs(all_scales_college_cor_load_splt[[1]])`) between the target SPLT parameters and the relevant subscales from the FSMI, K-SRQ, D&P, and UPPS-P questionnaires (adolescents were not administered the FSMI or D&P).
All point estimates for these correlations were between r = `r sprintf('[%0.2f, %0.2f]', scale_splt_cor_df_collegeminmax$min, scale_splt_cor_df_collegeminmax$max)`, with most confidence intervals not exceeding r = [-0.20, 0.20] (Figure \ref{fig:scalespltcorcollege}).
There is also no pattern with regard to the contrast and the scale: for example, the association between FSMI Mate-seeking and either $\epsilon_\text{dating}$ or the proportion of optimal responses dating contrasts (First or Last $1/2_{\text{dating}}$) is not noticeably stronger than the relation between these contrasts and FSMI Status.
Moreover, though the correlation between FSMI Mate-seeking and the First $1/2_{\text{dating}}$ contrast stands out, it's in the wrong direction.
This same absence of pattern is also apparent for the Popular/Unpopular - Hungry/Thirsty contrast.

```{r scalespltcorcollege, fig.width=5.875, fig.height=4, include=F}
print(scale_splt_cor_df_college_plot)
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/scalespltcorcollege-1.pdf}
\caption[SPLT parameter contrast correlations with latent variables, college]{SPLT parameter contrast correlations with latent variables, college sample. Subscripts for each parameter indicate the contrast condition versus baseline (Hungry/Thirsty). ``First 1/2'' and ``Last 1/2'' refer to contrasts in proportions of optimal responses in the first and last half of the run, respectively. U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire; - = Negative; + = Positive.}
\label{fig:scalespltcorcollege}
\end{figure}

Latent variable correlations with baseline (Hungry/Thirsty condition) performance also resulted in generally small estimates (r = `r sprintf('[%0.2f, %0.2f]', scale_splt_cor_df_collegeminmaxbl$min, scale_splt_cor_df_collegeminmaxbl$max)`), though some sensible patterns emerge (Figure \ref{fig:scalespltcorbaselinecollege}).
Specifically, overall performance in the last half of the run shows a small positive correlation with D&P Prestige, and UPPS-P Sensation Seeking and Perseverance, and small negative correlations with D&P Dominance, and UPPS-P Positive and Negative Urgency.
Examining correlations for the learning parameters, higher performance may be due to lower $\xi$ for those higher in Sensation Seeking, higher $\rho$ for those higher in Prestige, either or both lower $\xi$ and higher $\rho$ for those higher in Perseverance.
Lower performance may be the result of either or both lower $\rho$ and higher $\xi$ for those high in Dominance, and lower $\rho$ for those higher in Positive and Negative Urgency.

```{r scalespltcorbaselinecollege, fig.width=5.875, fig.height=2, include=F}
print(scale_splt_cor_df_baseline_college_plot)
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/scalespltcorbaselinecollege-1.pdf}
\caption[SPLT parameter baseline correlations with latent variables, college]{SPLT parameter baseline correlations with latent variables, college sample. ``First 1/2'' and ``Last 1/2'' refer to contrasts in proportions of optimal responses in the first and last half of the run, respectively. U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire; - = Negative; + = Positive.}
\label{fig:scalespltcorbaselinecollege}
\end{figure}

### Combined college and adolescent samples

```{r}
scale_splt_cor_df_combinedminmax <- dplyr::summarize(
  dplyr::filter(
    scale_splt_cor_df_combined,
    grepl('HngT_fac$', rhs),
    !grepl('^b_', rhs)), 
  min=min(est.std), max=max(est.std))

scale_splt_cor_df_combinedminmaxbl <- dplyr::summarize(
  dplyr::filter(
    scale_splt_cor_df_combined,
    grepl('HngryThrsty_fac$', rhs),
    !grepl('^b_', rhs)), 
  min=min(est.std), max=max(est.std))

scale_splt_cor_df_combinedN <- paste(c(
  paste0('N = ', sum(lavaan::inspect(all_scales_combined_cor_load_splt[[1]], 'nobs'))),
  paste0(sample_labels[lavaan::inspect(all_scales_combined_cor_load_splt[[1]], 'group.label')], 
         ' = ', 
         lavaan::inspect(all_scales_combined_cor_load_splt[[1]], 'nobs'))),
  collapse = ', ')
```

Next, using just the questionnaires administered to all four samples (`r scale_splt_cor_df_combinedN`), correlations were estimated between the target SPLT parameters and the relevant subscales from the K-SRQ, and UPPS-P.
The models that constrained correlations to be equal across samples fit better than those with correlations freed to differ between samples ($\Delta\text{AIC}_\text{max} = `r max(unlist(lapply(all_scales_combined_cor_load_splt[-4], AIC))-unlist(lapply(all_scales_combined_persamp_cor_load_splt[-4], AIC)))`$; models for the $b$ parameter did not converge).
All point estimates for these correlations remained small in the combined sample (r = `r sprintf('[%0.2f, %0.2f]', scale_splt_cor_df_combinedminmax$min, scale_splt_cor_df_combinedminmax$max)`), again, with most confidence intervals not exceeding r = [-0.20, 0.20] (Figure \ref{fig:scalespltcorcombined}).
None of the expected patterns appeared between (SPLT learning rate or performance) Dating contrasts and mate-seeking-relevant self-report scales, or Popular contrasts and status-relevant self-report.
However, there was an interesting consistency in positive correlations between $\rho$ for both contrasts and UPPS-P Sensation Seeking, and Positive and Negative Urgency, in line with the results from the college sample, above.
Although one might be tempted to link these constructs to reward sensitivity, it is clear in the baseline results below that this likely arises because of _negative_ correlation between these subscales and $\rho_{\text{baseline}}$.

```{r scalespltcorcombined, fig.width=5.875, fig.height=4, include=F}
print(scale_splt_cor_df_combined_plot)
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/scalespltcorcombined-1.pdf}
\caption[SPLT parameter contrast correlations with latent variables, all samples]{SPLT parameter contrast correlations with latent variables, all samples. Subscripts for each parameter indicate the contrast condition versus baseline (Hungry/Thirsty). ``First 1/2'' and ``Last 1/2'' refer to contrasts in proportions of optimal responses in the first and last half of the run, respectively. U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire; - = Negative; + = Positive.}
\label{fig:scalespltcorcombined}
\end{figure}

The latent variable correlations with baseline (Hungry/Thirsty condition) performance also resulted in generally small estimates (r = `r sprintf('[%0.2f, %0.2f]', scale_splt_cor_df_combinedminmaxbl$min, scale_splt_cor_df_combinedminmaxbl$max)`).
Both overall performance in the last half of the baseline condition, and the reward modifier, $\rho$, show negative correlations with UPPS-P Sensation Seeking, and Positive and Negative Urgency, as was the case in the college sample (Figure \ref{fig:scalespltcorbaselinecombined}).
The UPPS-P Perseverance variable continued to show somewhat attenuated correlations with $\xi$ (negative) and performance (positive).
Finally, though this does not seem to result in a concomitant correlation with performance, the three highly-correlated K-SRQ variables show small negative correlations with the noise parameter, $\xi$.

```{r scalespltcorbaselinecombined, fig.width=5.875, fig.height=2, include=F}
print(scale_splt_cor_df_baseline_combined_plot)
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/scalespltcorbaselinecombined-1.pdf}
\caption[SPLT parameter baseline correlations with latent variables, all samples]{SPLT parameter baseline correlations with latent variables, all samples. ``First 1/2'' and ``Last 1/2'' refer to contrasts in proportions of optimal responses in the first and last half of the run, respectively. U = Urgency, Premeditation, Perseverance, Sensation-seeking, \& Positive Urgency scale; K = Kids' Social Reward Questionnaire; - = Negative; + = Positive.}
\label{fig:scalespltcorbaselinecombined}
\end{figure}

As a final note, it should be clear from the plots above that if were one to have set an _a priori_ equivalence-testing threshold at $|r| = 0.20, \alpha = .05$, no correlation examined would be unambiguously different from the region of equivalence, and in many cases one would accept the null hypothesis that there is no correlation of any importance [@lakens2017].

## Self-report variable associations with behavioral outcomes

In order to check the concurrent validity of self-report measures of motives (and impulsivity/sensation seeking), associations with self-report behavior (and relationship status) were assessed.
Descriptive plots of responses to the following questions can be examined in the accompanying figures:

- "How many times, in the last six months, have you had sex/sexual relations," with sexual relations defines as "oral, vaginal or anal, and can be with male or female partners" (Figure \ref{fig:sexfreqplot})
- "With how many different people in the last 6 months" (Figure \ref{fig:sexfreqplot})
- "In the last 6 months, what percentage of the time when you've had sexual relations (oral, vaginal or anal) did you USE safe sex practices to prevent sexually transmitted diseases - condoms, blood tests, dental dams, etc.)." (Figure \ref{fig:useprotfreqplot})
- "In the past 30 days, on how many days did you have at least one drink of alcohol" (Figure \ref{fig:alcfreqplot})
- "In the past 30 days, on how many days did you have 5 or more drinks of alcohol in a row, that is, within a couple hours?" (Figure \ref{fig:alcfreqplot})

```{r sexfreqplot, fig.width=5.875, fig.height=3, include=F}
gridExtra::grid.arrange(sex_freq_plot, num_partnr_plot, sex_freq_plot_tds, 
                        num_partnr_plot_tds, nrow = 2, widths = c(4,3))
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/sexfreqplot-1.pdf}
\caption[Frequency of sex and number of partners, last six months]{Frequency of sex and number of partners, last six months.  Count is number of participants endorsing a particular response. LTR = Long term relationship, indicating that the participant reported they were in a such a relationship.}
\label{fig:sexfreqplot}
\end{figure}

```{r useprotfreqplot, fig.width=5.875, fig.height=2.5, include=F}
gridExtra::grid.arrange(useprot_freq_plot, useprot_plot_tds, nrow = 1, widths = c(2,1.1))
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/useprotfreqplot-1.pdf}
\caption[Frequency of safe sex practice, last six months]{Frequency of safe sex practice, last six months. Count is number of participants endorsing a particular response. LTR = Long term relationship, indicating that the participant reported they were in a such a relationship. Overlapping percentage ranges appear in the original questionnaire.}
\label{fig:useprotfreqplot}
\end{figure}

```{r alcfreqplot, fig.width=5.875, fig.height=3, include=F}
gridExtra::grid.arrange(alc_freq_plot, alc5_partnr_plot, alc_freq_plot_tds, 
                        alc5_partnr_plot_tds, nrow = 2, widths = c(4,3))
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/alcfreqplot-1.pdf}
\caption[Frequency of drinking, last 30 days]{Frequency of drinking, last 30 days. Count is number of participants endorsing a particular response.}
\label{fig:alcfreqplot}
\end{figure}

Overall, as would be expected, adolescent participants reported engaging in less sexual and drinking behavior than the college student participants. 
For those who were sexually active, the great majority also tended to report using safe sex practices.
College participants in long term relationships (adolescent participants were not asked) tended to have only one partner, and also tended to report higher numbers of sexual encounters.
Unpartnered college students were in the majority, tended to report a wider range of number of partners, and tended to more often endorse no, or lower numbers of sexual encounters.
Safe sex practices show similar distributions for partnered and unpartnered college students, with heavier use of the poles of the scale.
Most college students reported drinking at least one drink at least once in the last month, with a considerable number also reporting at least one heavy drinking day in the last month (five or more drinks in a session).
Given the low variability on these questions in the adolescent sample, associations with all measures proceeds by first examining the relation in the college samples, and then adding in the adolescent samples to assess sensitivity.

Finally, the difference in the distribution of the FSMI Mate-seeking items for those college students in long-term relationships is markedly different from those who are not (Figure \ref{fig:ltrfsmi}).
The item content for the FSMI Mate-seeking scale asks primarily about intentions to find new romantic and sexual partners. 
As such (and given social norms favoring monogamy in committed relationships), it is encouraging evidence for the validity of this measurement to see that those who are in long-term relationships report very low means on this scale, while those who are unpartnered have computed scores that span the full range.

```{r ltrfsmi, fig.width=3.25, fig.height=2, include=F}
print(ltr_fsmi)
```

\begin{figure}
\centering
\captionsetup{width=3.25in}
\includegraphics{figures/ltrfsmi-1.pdf}
\caption[FSMI Mate-seeking scale and long-term relationship status]{FSMI Mate-seeking scale and long-term relationship status. LTR = Long term relationship, indicating that the participant reported they were in a such a relationship. Values on X axis are binned means of six items.}
\label{fig:ltrfsmi}
\end{figure}

### Number of partners

The analyses in the next several sections test whether sexual behavior is associated with the FSMI Mate-seeking, or K-SRQ Sexual Relationships variables.
Initial models include both the focal motive variable and gender, which are compared to models in which the coefficient for the motive variable is fixed to zero.
Significant decrease in fit is taken as evidence for an association between the motive variable and the behavioral outcome.

In the second step, the initial model is augmented by adding UPPS-P Sensation Seeking, Positive Urgency, and Negative Urgency, with the same constraint of the focal motive variable tested.
Significant difference in fit here indicates that the focal motive is associated with the outcome even when conditioning on (or perhaps because of this conditioning) the UPPS-P impulsivity and sensation-seeking variables.

For the K-SRQ, an additional test is made by including the Admiration variable in the regression.
Significant decrease in fit when the coefficient for this variable is fixed to zero indicates that it has an association with the behavioral outcome over and above the variance it shares with the Sexual Relationships variable.
The sign of the coefficients between the unconstrained and constrained model will also be examined.

#### FSMI Mate-seeking

Constraining the coefficient of the FSMI Mate-seeking latent variable to zero in a model that regressed number of sexual partners on this variable and gender did not result in significantly worse fit when examining data from unpartnered college students (`r print_lrt_test_stats(fsmi_num_partners_lrt, 2, 3)`).
The coefficient for the effect of Mate-seeking on number of partners was, however, positive.

Next, we added to the above model the three UPPS-P variables (Sensation Seeking, Positive and Negative Urgency) and then compared this to a model that fixed the Mate-seeking coefficient to 0.
This did not result in a significantly worse fit (`r print_lrt_test_stats(uppsp_fsmi_partners_lrt2, 2, 2)`).
The coefficient for the Mate-seeking variable, when included, was positive.

#### K-SRQ Sexual Relationships

In a model with the K-SRQ Sexual Relationships (SR) and gender predicting number of partners, constraining the coefficient for SR to zero resulted in significantly worse fit both in the unpartnered college student sample (`r print_lrt_test_stats(ksrq_num_partners_null_lrt, 2, 4)`), and in the the full (college and adolescent) sample (which also includes age as a covariate; `r print_lrt_test_stats(ksrq_num_partners_null_all_lrt, 2, 6)`).
In both cases, the coefficient for SR is positive, indicating that those participants with higher levels on SR also report more sexual partners.

Comparing models with and without the K-SRQ Admiration variable included in the regression did not result in significantly worse fit in the unpartnered college sample (`r print_lrt_test_stats(ksrq_num_partners_admrtn_lrt, 2, 2)`), but did significantly reduce fit in the full samples (`r print_lrt_test_stats(ksrq_num_partners_admrtn_all_lrt, 2, 5)`).
In both cases, the coefficient for SR was positive and for Admiration was negative.
This indicates that for a given level on SR, a participant with a higher level of Admiration would be expected to have fewer partners.
Models that included K-SRQ Sociability did not converge, probably because of the very high correlation with the SR factor.

Removing SR from a regression that also included the three UPPS-P variables did not result in worse fit in the unpartnered college student sample (`r print_lrt_test_stats(uppsp_ksrq_num_partners_lrt)`), but did significantly reduce fit in the combined sample (`r print_lrt_test_stats(uppsp_ksrq_num_partners_all_lrt, 2, 6)`).
The coefficient for SR in these models was also positive.

### Number of sexual encounters

#### FSMI Mate-seeking

Comparing models with and without the FSMI Mate-seeking regression coefficient constrained to zero showed non-significant mis-fit in the unpartnered college sample (`r print_lrt_test_stats(fsmi_sfreq_lrt, 2, 2)`).
In the unconstrained model, the coefficient was negative and very close to zero.
Adding UPPS-P covariates to the model, and again testing whether constraining the Mate-seeking coefficient to zero also resulted in non-significant reduction in fit (`r print_lrt_test_stats(uppsp_fsmi_sfreq_lrt2, 2, 2)`), and in the unconstrained model the Mate-seeking coefficient remained negative.

#### K-SRQ Sexual Relationships

The test of K-SRQ SR showed non-significant reduction in fit when constraining its association with number of sexual encounters to zero in the unpartnered college sample (`r print_lrt_test_stats(ksrq_sex_freq_null_lrt,2,3)`), but significant reduction in fit in the full sample (`r print_lrt_test_stats(ksrq_sex_freq_null_all_lrt,2,6)`).
The coefficient for the SR variable in both cases was positive, indicating that those higher on SR tend to report having more sexual encounters (note that the coefficients for partnered and unpartnered college students were also both positive when considered separately).

Considering the models with and without the Admiration covariate showed non-significant difference in fit for the unpartnered college student sample (`r print_lrt_test_stats(ksrq_sex_freq_admrtn_lrt,2,2)`), as well as the full samples (`r print_lrt_test_stats(ksrq_sex_freq_admrtn_all_lrt,2,3)`).
As was the case with number of partners, the coefficients for the Admiration variable were negative (while SR remained positive), indicating that, given the same level of SR, those higher in Admiration tended to report fewer sexual encounters.

Not accounting for the association between number of sexual encounters and SR in a model that also included the three UPPS-P variables did not result in significant difference in fit in the unpartnered college sample (`r print_lrt_test_stats(uppsp_ksrq_sex_freq_lrt,2,2)`), but did result in a significant reduction in fit in the full sample (`r print_lrt_test_stats(uppsp_ksrq_sex_freq_all_lrt,2,3)`).
In both cases the coefficient for SR was positive.

### Safe sex practices

#### FSMI Mate-seeking

The fit difference tests for the effect of Mate-seeking on frequency of safe sex practices was not significant for either the unpartnered college sample (`r print_lrt_test_stats(fsmi_uprot_lrt,2,2)`) or for the full college sample (`r print_lrt_test_stats(fsmi_uprot_all_lrt,2,2)`).
The coefficients in both cases were very close to zero.
Including the three UPPS-P variables as covariates also resulted in no significant fit decrease when constraining the effect of Mate-seeking to be zero in the unpartnered (`r print_lrt_test_stats(uppsp_fsmi_uprot_lrt2,2,2)`) and combined college student sample (`r print_lrt_test_stats(uppsp_fsmi_uprot_all_lrt2,2,2)`).

#### K-SRQ Sexual Relationships

Neither the unpartnered college student sample (`r print_lrt_test_stats(ksrq_use_protection_null_lrt,2,2)`), nor the full sample 
(`r print_lrt_test_stats(ksrq_use_protection_null_all_lrt,2,2)`) showed significant decrease in fit when constraining the effect of the SR variable on frequency of safe sex practice to zero.
The coefficient in both cases was positive, indicating that those higher on SR also tend to report more frequent practice of safe sex.

In a model that also included Admiration as a covariate, removing this variable from the regression did not reduce fit significantly for ether unpartnered college students (`r print_lrt_test_stats(ksrq_use_protection_admrtn_lrt,2,2)`) or the full sample (`r print_lrt_test_stats(ksrq_use_protection_admrtn_all_lrt,2,2)`).
The coefficients were positive for both SR and Admiration in these models, though close to zero.
Models with the three UPPS-P variables included demonstrated the same pattern for the unpartnered college students (`r print_lrt_test_stats(uppsp_ksrq_use_protection_lrt,2,2)`) and combined sample (`r print_lrt_test_stats(uppsp_ksrq_use_protection_all_lrt,2,2)`).

### Number of drinking days

The following models test whether drinking behavior is associated with the FSMI Status, or K-SRQ Admiration variables.
As above, initial models include both the focal motive variable and gender, which are compared to models in which the coefficient for the motive variable is fixed to zero.
Significant decrease in fit is taken as evidence for an association between the motive variable and the behavioral outcome.

In the next step, the initial model is augmented by adding UPPS-P Sensation Seeking, Positive Urgency, and Negative Urgency, with the same constraint of the focal motive variable tested.
Significant difference in fit here indicates that the focal motive is associated with the outcome even when conditioning on (or perhaps because of this conditioning) the UPPS-P impulsivity and sensation-seeking variables.

For the K-SRQ, an additional test is made by including the Sociability variable in the regression.
Significant decrease in fit when the coefficient for this variable is fixed to zero indicates that it has an association with the behavioral outcome over and above the variance it shares with the Admiration variable.
The coefficients between the unconstrained and constrained model will also be examined.

#### FSMI Status

In the college sample, constraining FSMI Status to have no association with number of drinking days did not significantly reduce model fit (`r print_lrt_test_stats(fsmi_alcdays_lrt,2,2)`), with a coefficient for the effect, when it was freed, close to zero.
This remained the case whether or not the UPPS-P variables were included in the model (`r print_lrt_test_stats(uppsp_fsmi_alcdays_lrt2,2,2)`).

#### K-SRQ Admiration

Constraining the Admiration variable to zero in the regression of number of drinking days in the college sample resulted in a significant reduction in fit (`r print_lrt_test_stats(ksrq_alc_days_null_lrt,2,4)`).
This was also the case for the full sample (`r print_lrt_test_stats(ksrq_alc_days_null_all_lrt,2,5)`).
In both cases, the freed coefficient was positive, indicating that those with higher values on Admiration also reported more days where they had at least one drink (in the past 30 days).

The exclusion of the Sociability variable also led to significant decreases in fit for both the college sample (`r print_lrt_test_stats(ksrq_alc_days_admrtn_lrt,2,4)`), as well as the full sample (`r print_lrt_test_stats(ksrq_alc_days_admrtn_all_lrt,2,6)`).
In both cases, the freed coefficient for Sociability was positive, while the coefficient for Admiration became negative (with increased standard error).
This indicates that participants reporting higher sociability also tend to report more drinking days, and that for a given level of Sociability, higher Admiration may be associated with fewer drinking days.

When tested in a model that included the three UPPS-P variables, constraining the Admiration effect on drinking days to zero did not reduce fit in the college sample (`r print_lrt_test_stats(uppsp_ksrq_alc_days_lrt,2,2)`) or the combined sample (`r print_lrt_test_stats(uppsp_ksrq_alc_days_all_lrt,2,2)`).
The coefficient for the association between Admiration and drinking days remained positive in these models.

### Number of heavy-drinking days

#### FSMI Status

We did not reject the null of no association between FSMI Status and number of heavy drinking days in either the initial model (`r print_lrt_test_stats(fsmi_alc_5drnk_lrt,2,2)`) or the model including the UPPS-P variables (`r print_lrt_test_stats(uppsp_fsmi_alc5drk_lrt2,2,2)`).
The coefficient of the association was positive in the first case, and close to zero (very slightly negative) in the UPPS-P model.

#### K-SRQ Admiration

The results for the the K-SRQ Admiration association with number of _heavy_ drinking days were much the same as those for number of days having consumed any alcohol at all.
There was a significant decrease in fit when the coefficient for Admiration was set to zero for both the college (`r print_lrt_test_stats(ksrq_alc_5drnk_null_lrt,2,4)`) and full (`r print_lrt_test_stats(ksrq_alc_5drnk_null_all_lrt,2,5)`) samples, with positive coefficients for the freed parameters.
In models including the association with Sociability, significant fit differences were seen when fixing that association to zero for both the college (`r print_lrt_test_stats(ksrq_alc_5drnk_admrtn_lrt,2,3)`) and full sample (`r print_lrt_test_stats(ksrq_alc_5drnk_admrtn_all_lrt,2,6)`). In both cases, freeing the Sociability coefficient resulted in a negative Admiration coefficient with smaller magnitude than the models excluding Sociability.
In addition, in models including the UPPS-P variables, the coefficient for Admiration could be constrained to zero without significant fit difference in the college (`r print_lrt_test_stats(uppsp_ksrq_alc_5drnk_lrt,2,2)`) and combined sample (`r print_lrt_test_stats(uppsp_ksrq_alc_5drnk_all_lrt,2,2)`).

## SPLT relation to outcomes

Some models regressing the frequency of safe sex practices on the SPLT parameter $b$ did not converge, so comparisons involving these models were not tested.
Since this parameter was not the focus of the investigation, these convergence issues were not resolved.
No model comparisons met the $\alpha = .005$ threshold, and the direction of coefficients were often not in the expected direction -- for example, the association between number of partners and the $\epsilon$ Dating/Looking contrast is negative (Table \ref{tab:spltbehavchisq}).

\begin{table}
\centering
\begin{threeparttable}
\caption{SES and YRBS behavior reports regressed on SPLT variables} \label{tab:spltbehavchisq}
\footnotesize
```{r spltbehavchisq, results='asis'}
knitr::kable(splt_sems_df_4table[,-(1:2)],
             col.names = c('Behavior variable',
                           'SPLT variable',
                           'Coef sign$_{\\text{ltd}}$',
                           '$\\Delta\\chi_{\\text{ltd}}^{2}$',
                           '$p_{\\text{ltd}}$',
                           'Coef sign$_{\\text{full}}$',
                           '$\\Delta\\chi_{\\text{full}}^{2}$',
                           '$p_{\\text{full}}$'),
             align = c('l', 'l', 'r', 'r', 'r', 'r', 'r', 'r'),
             booktabs = T, format = 'latex', escape = F,
             linesep = '')
```
\begin{tablenotes}
\footnotesize
\item First and Last 1/2 $p_{\text{opt}}$ refer to the proportion of optimal responses made during the first and last half of the SPLT run, respectively. SPLT variable contrasts are between the indicated condition and the minimally social Hungry/Thirsty condition.
\item The ``ltd'' and ``full'' subscripts refer to the samples used in model fitting. For the sexual behavior questions, the limited (ltd) sample includes just unpartnered participants from the two college samples, and for the alcohol use questions, the ltd sample includes all participants from the two college samples. The full sample includes both college and both adolescent samples. ``Coef sign'' indicates the sign of the coefficient of the association between the behavior variable and SPLT contrast variable.
\end{tablenotes}
\end{threeparttable}
\end{table}


# Discussion, Aim 2

In these analyses, we observed very little evidence that behavior on the Social Probabilistic Learning Task (SPLT) is associated with either self-reported traits or health-related behavior.
There was relatively adequate evidence that the self-report trait scales successfully measured the relevant constructs, eliminating this as a possible explanation for the absence of strong associations with SPLT behavior.
Some measurement issues, specifically poor measurement-model fit and high correlations among K-SRQ scales, qualify these observations.

## Self-report scale measurement and associations

All confirmatory factor analyses of the self-report measurement models showed questionable fit, with the exception of the FSMI Mate-seeking and Status model.
Poor fit may be due to incorrect specification of the relation of the items to latent constructs, such as assuming the wrong number of factors, or an incorrect latent variable hierarchy. 
Factor structure was not tested as part of these analyses, but there is some evidence in the correlations between latent variables of considerable overlap that may warrant consideration of alternative structures.
For example, Negative and Positive Urgency in the UPPS-P may not be separable factors, or may be lower-order factors of a higher-order impulsivity variable [@cyders2014].
Poor fit and high subscale correlations for the K-SRQ scales, which only has forward keyed items, may also simply be the result of acquiescence bias.
Extreme responding driven either by social desirability or miscalibrated item content (e.g., almost no one dislikes being looked up to, or kissing) is also a concern with this scale, which shows skew in responses to almost every item.
Indeed, two subscales were not included in these analyses because extreme responding in one direction was so pronounced.
When measurement non-invariance was observed, sensitivity tests indicated without exception that substantive results were not sensitive to them.

Despite possible problems with measurement, correlations between latent variables, associations with age and puberty, and associations with self-report behavior, showed patterns that may increase our confidence in the validity of these self report instruments.
Motivation to seek out new romantic and sexual partners, and enjoyment of romantic and sexual behaviors, should have a fairly direct influence on self-reported behaviors like number of sexual partners and frequency of sexual behavior.
The number of sexual partners was consistently significantly associated with the K-SRQ Sexual Relationship measure of socio-sexual reward in both the set of college participants not in a long-term relationship (unpartnered) and in the full sample.
This association was somewhat attenuated when including the UPPS-P variables in the unpartnered college student sample.
Consistent but less robust positive associations were also seen between the FSMI Mate-seeking measure and number of sexual partners.

The K-SRQ Sexual Relationships measure also showed consistent positive associations with frequency of sexual encounters, whether or not UPPS-P variables were included, but significant reduction in model fit was only observed when looking at the full sample.
Interestingly, the association between number of recent sexual encounters and mate-seeking, was close to zero or even slightly negative.

Neither the FSMI Mate-seeking nor K-SRQ Sexual Relationships scale showed a robust association with frequency of safe sex practice, regardless of the sample analyzed.
The K-SRQ coefficients were positive, possibly indicating a slight tendency for those who seek social reward from sexual and romantic experiences to also be more well informed about best practices for reducing risk.
The lack of association between Mate-seeking and safe sex practice seems to be inconsistent with findings during scale development that, among single individuals, Mate-seeking was positively correlated with condom use [@neel2015].
However, this finding was also relatively weak, and was more focused on one particular aspect of safe sex practice, whereas the current investigation uses a more diffuse measure that incorporates more subjective information about a more full range of safe sex practice.

Several other findings also support the validity of the measures of mate-seeking motive.
Helpfully for the validation of FSMI Mate-seeking, and as was first reported by @neel2015, this measure was directly impacted by current relationship status, with those in long-term-relationships scoring much lower.
The K-SRQ Sexual Relationships scale also shows an expected positive association with both age and puberty, though this is somewhat more pronounced in the subset of female participants compared to male participants.
The two sexual motivation subscales also show a positive correlation, though it is somewhat small.
Overall, this pattern of associations between the measures of motives, self-report sexual behavior, and relationship status is consistent with our expectations of these constructs, and helps support their validity.

However, many of the correlations between K-SRQ Sexual Relationships and other measures are rather high in unexpected ways, possibly indicating less clarity in its interpretation.
For example, it correlates more strongly with FSMI Status, and Prestige strategy than with FSMI Mate-seeking.
As mentioned above, it correlates extremely highly with both K-SRQ Admiration and K-SRQ Sociability.

The several scales measuring aspects of status motivation, FSMI Status, Prestige, and K-SRQ Admiration all showed strong correlations.
Dominance strategy correlated strongly with FSMI Status but not with either Prestige or K-SRQ Admiration.
Given that dominance is a distinct and strategy from prestige, the low correlation between these variables simply indicates that one might adopt either or both strategies in pursuit of achievement of social rank, consistent with the original report of @cheng2013.
The high correlation of Prestige with K-SRQ Admiration (and high correlation of both with FSMI Status) indicates that the Admiration scale is perhaps primarily measuring individual differences in enjoyment of receipt of prestige-based status in particular.
This is also suggested both by the scale names and overlap in item content (e.g., Prestige: "I am held in high esteem by those I know"; K-SRQ: "I like achieving recognition from others").
The modest correlation between Admiration and UPPS-P Perseverance, which is very similar between both the college and full sample, is consistent with this interpretation that Admiration is capturing status motivation.

Status motives as measured in the college student sample using the FSMI Status scale did not show significant associations with number of drinking days or number of heavy drinking days, and did not show any consistent pattern in the sign of the coefficients.
Admiration reward sensitivity was significantly positively associated with both number of drinking days and heavy drinking days.
However, Sociability also showed a significant relation even when controlling for Admiration, which showed a _negative_ association with both drinking outcomes when Sociability was included.
This is actually somewhat consistent with the investigation by @allen2005, which found that during adolescence, popularity was both associated with well-adjusted development but also increased exposure to peer groups where minor deviant behavior was socialized.
Though alcohol consumption is almost certainly considered less (or not at all) deviant among the college participants, a similar principal may apply here.
It is possible that sensitivity to admiration from peers is related to alcohol use just because enjoying admiration, to some extent, requires enjoying socializing as well.
When taking into account motivation to socialize, specifically, concerns about receiving admiration from peers seem to be associated with more cautious consumption. 

To use these self-report questionnaires to test whether individual differences in SPLT behavior arise from individual differences in related motives, it is essential that the self-report measures be valid instruments for those motives.
It must also be the case that the way those motives are elicited in the task, specifically by presenting information about dating-pool availability or social status, is related to the way the constructs are conceptualized on the item level.
In other words, does it make sense that someone who scores high on either the FSMI Mate-seeking or K-SRQ Sexual Relationships scale would care about whether another person was currently in a relationship or looking for someone to date?
Would someone who is seeking social status, or who enjoys receiving recognition from others care about whether a peer is popular or unpopular?
Of course, it is difficult to say for certain, but in the case of the mate-seeking motive questionnaires, the answer seems to be obviously yes.
With regard to the status-motive questionnaires, the answer is less clear, and may depend on whether a person's strategy for achieving status requires knowledge of their peers status.
Just as there are at least two strategies for gaining status, there are also seemingly two sub-types of popular peers -- those who are popular but not necessarily well-liked, and those who are popular _because_ they are well-liked [@debruyn2006].
While there is some evidence that being friends with popular peers, per se, confers status [@meuwese_friends_2017], it is not clear whether this depends on the strategies of these particular peers.
Based on the indications of construct validity reviewed above, and a plausible case for the relevance of the task manipulations to status and mate-seeking motives, these self-report scales are likely to function as adequate criteria for the relevance of individual differences in task behavior to motivation in these domains.

## SPLT variable associations

Correlations between the self-report latent variables and SPLT variables were all very low.
While the contrasted-SPLT variable correlation estimates did not show readily interpretable patterns, there was some indication that correlations with the baseline (Hungry/Thirsty condition) variables did show a sensible pattern in their sign.
This provides some evidence that individual differences in learning behavior, overall, is associated with self report trait differences, but that differences between learning conditions are either not related to individuals' trait differences, or that the task measurement is not sensitive enough to detect these relations.
Though the largest associations are discussed for the sake of completeness, note well that these effects are small, to very small, and arise out of a large number of estimates.

In the combined sample, baseline learning behavior seemed to be weakly associated with several subscales of the UPPS-P.
Impulsivity, as measured by the Urgency scales, showed negative relations with performance during the last half of the run, while the Perseverance scale, a measure of ability to stay focused, was positively associated with performance.
Interestingly, these same individual differences are also weakly associated with different model parameters, suggesting that if there is a relation between these traits and performance, they may arise through different psychological mechanisms.
Specifically, poorer performance for those higher in Urgency may be due to lower sensitivity to any particular piece of feedback, as indexed by the parameter $\rho$.
Perseverance, on the other hand, is also weakly positively associated with lower $\xi$, indicating that greater ability to focus may reduce the amount of random (or erroneous) responding, thereby increasing performance.
Though these effects are very weak, and possibly false-positives, their sign and association with both performance and particular model parameters is coherent enough to be suggestive. 

When the college sample was isolated, we observed similar patterns with the addition of both weak negative associations between performance and Dominance strategy, and weak positive associations between performance and Prestige and sensation seeking.
Prestige and Dominance associations with performance seemed to be the result of opposite patterns of, respectively, higher and lower sensitivity to reward ($\rho$), and lower and higher random responding ($\xi$).
On the other hand, the relations of Sensation Seeking to performance may be explained primarily by its negative correlation with $\xi$.
This Prestige correlation is consistent with the view that such a strategy requires individuals to demonstrate competency, and though rather contrived, this task is an opportunity to perform well.
Dominance-related popularity strategies have been associated with relational aggression and academic disengagement [@debruyn2006;@kiefer2008], which could explain this negative relation to performance, possibly through decreased focus.
If there is a true relation between task performance and Sensation Seeking, it is much less straightforward -- while some authors have suggested that purported heightened reward sensitivity during adolescence could lead to improved learning [@davidow2016;@mccormick2017], Sensation Seeking in these data correlate more strongly with the noise parameter, $\xi$, than with reward sensitivity, $\rho$.
Generally, these patterns of correlations with baseline task behavior highlight one of the potential strengths of this method (which would be more easily verified with higher precision), which is that it allows decomposition of behavior into distinct psychological processes that may be difficult to disentangle using other means [@vandenbos2017].

Although correlations between SPLT contrast variables and self-report scales are of a similar magnitude to the baseline correlations, they have a much less straightforward interpretation.
For example, in the combined samples, college-only samples, higher sensation seeking and impulsivity as measured by UPPS-P subscales are weakly positively correlated with the $\rho$ parameter differences between both the Dating and Popular conditions versus baseline, but much more weakly, still, with overall performance.
Only when the combined sample is considered does Passivity (which measures enjoyment of being told what to do by others) correlate negatively with final performance differences in the Dating condition, while perseverance correlates positively with early performance in the dating condition.
These results do not seem to be consistent with the expectations of this investigation or with literature about performance effects within these specific domains.
Perhaps they may be of use to a future meta-researcher.

Finally, the SPLT variables also showed no coherent associations to any of the self-reported behavioral outcomes.
These analyses were included here primarily to further validate task behavior as a measure of social motives in the case that SPLT behavior and self-report scales showed correspondence, and to test whether SPLT behavior might be a potentially _better_ predictor of health-related behaviors than self-report trait measures.
In the context of the results discussed above, the lack of significant associations between SPLT variables and behavioral outcomes serves as another indication that the task either does not capture individual differences in relevant social motives, or does so only very imprecisely.

## Conclusions

Individual differences in learning performance and model-estimated variables derived from the SPLT do not clearly measure individual differences in social motives, nor other potentially related constructs.
Self-report trait measures, on the other hand, were observed to relate coherently with each other, and with selected health-related behaviors, though there were also some indications that they could be further refined to hone their effectiveness.

