---
documentclass: apa6
classoption: "margin=1in,man,floatsintext"
header-includes:
  # - \usepackage[para,online,flushleft]{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage[singlelinecheck=false]{caption}
  - \captionsetup[subfigure]{singlelinecheck=on, labelfont=normalfont}
  - \captionsetup[figure]{labelfont=it}
  - \shorttitle{SOCIAL MOTIVES \& REINFORCEMENT LEARNING}
title: "Measurement of social motives: relations between self report and reinforcement learning"
author: "John C. Flournoy"
output: 
  pdf_document:
    keep_tex: yes
    template: "default-1.17.0.2_nogeom.tex"
bibliography: "/home/jflournoy/code_new/social-motives-rl-writeup/dissertation.bib"
csl: "/home/jflournoy/Rlibs/probly/bib/apa-old-doi-prefix.csl"
---

```{r setupch4, include=FALSE}
knitr::opts_chunk$set(fig.path = 'figures/', echo = F, warning = F, error = F, message = F)
library(probly)
load('rda/fsmi-cfa.rda')
ggplot2::theme_set(ggplot2::theme_minimal())

sample_abbreviation_note <- 'FCA: foster-care-involved adolescents; CA: community adolescents; CSYA: college students; CSYA-O: college students, online.'
```

# Background

- Review Neel, fsmi lit
  - Evidence for validity
  - Relation to outcomes
    - note expected long-term-relationship differences
- Review Foulkes, SRQ lit
  - Evidence for validity
  - Relation to outcomes
    - note there are not expected long-term-relationship differences
- Review sensation seeking
  - Validity
  - Relation to outcomes
- Describe outcomes of interest
  - FSMI Mate-seeking (FSMI-MS) and K-SRQ sexual relationship (K-SRQ-SR)
    - Normative sexual behavior
      - Number of partners
      - Number of sexual encounters
  - FSMI-MS, K-SRQ-SR, FSMI status, K-SRQ admiration, Sensation seeking
      - Safe sex practice
      - Substance use and sex
      - Alcohol use
  
# Method

## Scale validation

- Confirmatory factor analysis procedure
- Correlations between latent factors on FSMI, K-SRQ, UPPS-P
- Relation of sexual behavior self report to FSMI-MS, K-SRQ-SR
  - note long term-relationship differences
- Age and puberty associations with FSMI, K-SRQ, UPPS-P

## Self report relation to SPLT

- Using raw $P_{O}$ calculated from behavior
- Using model-estimated learning parameters

## Self report relation to outcomes

- Regressing ordinal SES and YRBS scales on latent factors
- Controling for age and gender
- Test sensation seeking against motives-specific scales

## SPLT relation to outcomes

- Regressing ordinal SES and YRBS scales on learning performace
- Regressing ordinal SES and YRBS scales on model parameters

_Note on planned moderation analyses:_ I had planned to use the CARE-R as a moderating variable between risk behaviors and motives. I'm deprioritizing these analyses for now.



# Results

## Scale validation

### FSMI CFA

- College sample only
- Mate-seeking and Status scales
  - Fit, Intercorrelations
    - A CFA for mate-seeking and status factors fit well (N = `r lavaan::inspect(fsmi_cfa_matestat, what = 'ntotal')`, RMSEA = `r round(fsmi_matestat_cfa_fit[[1]][['rmsea']],3)`, $\hat\gamma$ = `r round(fsmi_matestat_cfa_fit[[2]][['gammaHat']],3)`), with the two factors showing a small, positive, non-significant correlation ($r = `r round(fsmi_mate_stat_cor[['std.all']],2)`, Z = `r round(fsmi_mate_stat_cor[['z']],2)`$).
    - Note that item 54, "I do not worry very much about losing status" has a particularly low loading on the status factor (standardized loading = `r round(fsmi_mate_stat_par[fsmi_mate_stat_par$op == '=~' & fsmi_mate_stat_par$rhs == 'fsmi_qs_54','std.all'],2)`).
  - No age/puberty (too narrow a range)

### K-SRQ CFA

- College and adolescent samples
- Note that negative social potency and prosocial iteractions items were very highly skewed, so these subscales were dropped before any further analysis was performed.

- This leaves four subscales: admiration, passivity, sexual relationshups, and sociability

A CFA for the four K-SRQ factors was tested for metric invariance across samples by comparing a model with loadings constrained to equality to one with loadings free to vary across the four groups.
This resulted in considerably worse model fit as determined by a change in the MFI > .01.
Loading constraints were removed from four items ("I like it if others looks up to me", "I like being a member of a group/club", "I like kissing", "I like flirting") that load onto Admiration, Sociability, and  Sexual Relationships, respectively.
This resulted in adequately small fit differences between the unconstrained and invariant models (`r ksrq_cfa_metric_invar_stats`).
The final model showed somewhat poor fit ($\text{N}_{\text{total}}$ = `r lavaan::inspect(ksrq_cfa, what = 'ntotal')`, RMSEA = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['rmsea']],2)`, MFI = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['mfi']],2)`, $\hat\gamma$ = `r round(ksrq_cfa_fit$ksrq_cfa_metric[['gammaHat']],3)`). 

Models with and without constraintes between latent variances and covariances were compared using the Akaike information criterion (AIC).
Constraining equivalent latent covariance structures across groups resulted in lower AIC ($\Delta\text{AIC} = `r round(ksrq_cfa_fit$delta2[['AIC']])`$).
Loadings without constraints across groups do not contribute to this comparison, so the magnitude of correlations was compared across the metric invariant and unconstrained models (while mainting the constraint of covariance across groups).
The differences in correlations were in the range _r_ = [`r round(min(ksrq_cor_diff),2)`, `r round(max(ksrq_cor_diff),2)`], indicating very low sensitivity the observed measurement invariance.
To maintain content coverage [@borsboom2006], the unconstrained model is used, though sensitivity to constraints on the four non-invariant loadings is assessed again.
Correlations among the Admiration, Sociability, and  Sexual Relationships subscales were all high (_r_ = [.67,.84]). 
Correlations of Passivity with these three subscales was generally low, but positive (_r_ = [.07, .21]; see Table ??? for all latent variable correlations).
This pattern of correlation is similar to what was seen in the scale construction sample [@foulkes2014].

Variation in latent factors was described by a quadratic age regression, with age effects allowed to vary by gender.
The partial metric invariance model was used to test whether constraining regression coefficients across samples resulted in a better fitting model as assessed by the AIC.
Constraining the regression coefficients across sample results in reduced AIC ($\Delta\text{AIC} = `r round(diff(ksrq_cfa_metric_agesem_fit[,'AIC']),0)`$), indicated better model fit.
Both standardized coefficients were extremely similar between the model with partial metric invariance and the model with all loadings constrained to equality across samples ($\Delta\beta = `r sprintf('[%0.2f, %0.2f]', min(ksrq_agesem_betadiffs), max(ksrq_agesem_betadiffs))`$), so the fully constrained model is interpretted.
Model expected trend lines show somewhat higher scores in late adolescent participants (Figure \ref{fig:ksrqagepredplot}), with the largest quadratic effects (averaged across gender) for Admiration, Socioability, and Sexual Relationships (Table \ref{tab:ksrqage}).
Gender differences in the linear model coeficients were fairly small.

```{r ksrqagepredplot, fig.width=5.875, fig.height=4, include=F}
ksrq_age_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/ksrqagepredplot-1.pdf}
\caption[K-SRQ model-predicted age associations]{K-SRQ model-predicted age associations Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:ksrqagepredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{K-SRQ factors regressed on age and gender} \label{tab:ksrqage}
\small
```{r lowfidev, results='asis'}
knitr::kable(ksrq_std_par_table,
             col.names = c('K-SRQ factor', 'Sample', 
                           '$\\text{Age}$', '$\\text{Age}\\times\\text{Gender}$', 
                           '$\\text{Age}^2$', '$\\text{Age}^2\\times\\text{Gender}$'),
             booktabs = TRUE, format = 'latex', escape = F)
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}

We also described variation in latent factors by a quadratic pubertal development scale (PDS) regression, with effects allowed to vary by gender.
The partial metric invariance model was used to test whether constraining regression coefficients across samples resulted in a better fitting model as assessed by the AIC.
Constraining the regression coefficients across sample results in reduced AIC ($\Delta\text{AIC} = `r round(diff(ksrq_cfa_metric_pdssem_fit[,'AIC']),0)`$), indicated better model fit.
Both standardized coefficients were extremely similar between the model with partial metric invariance and the model with all loadings constrained to equality across samples ($\Delta\beta = `r sprintf('[%0.2f, %0.2f]', min(ksrq_pdssem_betadiffs), max(ksrq_pdssem_betadiffs))`$), so the fully constrained model is interpretted.
Model expected trend lines show somewhat lower scores associated with lower PDS, primarily for female participants (Figure \ref{fig:ksrqpdspredplot}), with the largest linear effects (averaged across gender) for Socioability, and Sexual Relationships (Table \ref{tab:ksrqpds}).
Though a model including the quadratic effect of PDS was estimated, this was done to allow deceleration in the linear trend to approximate the observed asymptotic relation.

```{r ksrqpdspredplot, fig.width=5.875, fig.height=4, include=F}
ksrq_pds_pred_plot
```

\begin{figure}
\centering
\captionsetup{width=5.875in}
\includegraphics{figures/ksrqpdspredplot-1.pdf}
\caption[K-SRQ model-predicted PDS associations]{K-SRQ model-predicted age associations. Best fit lines are expected latent variable values drawn from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples. Best fit lines are plotted separately for each sample, but because regression coefficients are identical, and the expected trajectories are otherwise very similar, they are not separately labeled. Points are individual, model predicted, latent variable values.}
\label{fig:ksrqpdspredplot}
\end{figure}

\begin{table}
\centering
\begin{threeparttable}
\caption{K-SRQ factors regressed on age and gender} \label{tab:ksrqpds}
\small
```{r ksrqpds, results='asis'}
knitr::kable(ksrq_std_pds_par_table,
             col.names = c('K-SRQ factor', 'Sample', 
                           '$\\text{PDS}$', '$\\text{PDS}\\times\\text{Gender}$', 
                           '$\\text{PDS}^2$', '$\\text{PDS}^2\\times\\text{Gender}$'),
             booktabs = TRUE,
             escape = F, format = 'latex')
```
\begin{tablenotes}
\footnotesize
\item `r sample_abbreviation_note`
\item Values are standardized regression coefficients from the model with all loadings, latent variances and covariances, and regression coefficients constrained to be the same across samples, with 95% confidence intervals in brackets.
\end{tablenotes}
\end{threeparttable}
\end{table}
  
### UPPS-P CFA

- College and adolescent samples
- Fit
- Intercorrelations
- Age/puberty

### All scale correlations

- Within college sample, separated by +/- long-term-relationship status
- College + adolescents (K-SRQ, UPPS-P)
  - not separated by LTR (no data for adolescents)
- all-item EFA?
- Dominance and prestige scale relations to FSMI status, and K-SRQ admiration & passivity

### Relation to normative sexual behavior

This serves to ensure that the the Fundamental Social Motives Inventory, Mate-Seeking (FSMI-MS), and Kids' Social Reward Questionnaire (K-SRQ-SR) scales are measuring what they ought to be. 
We would expect positive relations between FSMI-MS and number of sexual partners in people who are not in a long term relationship (LTR-). 
For those in a long term relationship (LTR+), the reported number of sexual partners is strongly modally 1, and the average score on mate-seeking is very low.
This is the expected behavior of the scale, which is supposed to measure a persons present desire to meet new romantic or sexual partners [@neel2015].
The K-SRQ-SR measures a person's enjoyment of socio-sexual behavior, with items asking about kissing, flirting, and having crushes.
This is in fact a modified version of the original SRQ-SR which instead featured questions about the enjoyment more specifically of sexual activity [@foulkes2014]. 

_Note gender and age controls used where appropriate_

#### FSMI
- Number of partners
  - FSMI mate-seeking +related to num partners 
    - Within LTR- group only, $N = 157, \beta = .24$
    - _Also_ across all college participants, $N = 214, \beta = .24$
  - Women who are not in long term relationships report fewer partners than men ($\beta = -.51$)
  - No age control (too narrow a range)
- Number of encounters
  - Group analysis in partnered and unpartnered college students
  - Close to 0 FSMI mate-seeking relation with number of encounters 
    - $N = 203; \beta_{\text{LTR-}} = -.09; \beta_{\text{LTR+}} = -.03$
  - Gender related just for LTR- Women LTR- have fewer encounters than men
    - $\beta_{\text{LTR-}} = -.47; \beta_{\text{LTR+}} = .00$; gender-effect difference not tested
  - No age control (too narrow a range)

#### K-SRQ

- Number of partners
  - K-SRQ sexual relationships +related to num partners
    - Within LTR- college sample only, $N = 140, \beta = .35$
      - Controlling for gender, $\beta_{\text{F-M}} = -.19$
    - _Also_ across all participants that have data, $N = 260, \beta = .34$
      - Controlling for age and gender, $\beta_{\text{age}} = .42, \beta_{\text{F-M}} = -.03$
- Number of encounters
  - Group analysis in partnered and unpartnered college students (N = 203)
    - K-SRQ sexual relations scale association in LTR- group
      - $\beta_{\text{LTR-}} = .35; \beta_{\text{LTR+}} = .13$
    - Gender effect not significant but expected direction
      - Women LTR- have fewer encounters than men, $\beta_{\text{LTR-}} = -.14; \beta_{\text{LTR+}} = .08$; gender-effect difference not tested
    - No age control (too narrow a range)
  - Across all college and adolescent participants with data (N = 253)
    - K-SRQ sexual relations scale $\beta=.35$
    - Controlling for age and gender ($\beta_{\text{age}}=.49, \beta_{\text{F-M}} = .06$)

#### FSMI-MS & K-SRQ-SR

- Unpartnered college students (N = 132)
- Number of partners: $\beta_{\text{FSMI-MS}}=.18,\beta_{\text{K-SRQ-SR}}=.30,\beta_{\text{F-M}}=-.18$
- Number of encounters: $\beta_{\text{FSMI-MS}}=-.24,\beta_{\text{K-SRQ-SR}}=.39,\beta_{\text{F-M}}=-.15$

## Self report relation to SPLT

- Correlations between 
  - SPLT 
    - $P_{O}$ = proportion of optimal presses in last half of condition's trials
      - Contrasted DL - HT, PU - HT
    - parameters $\epsilon, \rho, \xi$
      - Contrasted transformed values (DL - HT, PU - HT)
  - All scales 
    - FSMI MS & Status
    - K-SRQ SR, admiration, passivity, sociability
    - UPPS-P

## Self report relation to outcomes

- FSMI-MS, K-SRQ-SR, FSMI status, K-SRQ admiration, Sensation seeking
  - For each compare to model with and without sensation seeking/impulsivity
    - Safe sex practice
      - "In the last 6 months, what percentage of the time when you've had sexual relations (oral, vaginal or anal) did you USE safe sex practices to prevent sexually transmitted diseases - condoms, blood tests, dental dams, etc.)"
    - Substance use and sex
      - "Did you drink alcohol or use drugs before you had sexual intercourse the last time you had sex?"
    - Alcohol use
      - "In the past 30 days, on how many days did you have at least one drink of alcohol?"
      - "In the past 30 days, on how many days did you have 5 or more drinks of alcohol in a row, that is, within a couple hours?"

## SPLT relation to outcomes

This section may be largely perfunctory if in fact the task is not sensitive to individual differences in motives.

- Contrasted model parameters and mean optimal behavior as regressors for 
  - Safe sex practice
    - "In the last 6 months, what percentage of the time when you've had sexual relations (oral, vaginal or anal) did you USE safe sex practices to prevent sexually transmitted diseases - condoms, blood tests, dental dams, etc.)"
  - Substance use and sex
    - "Did you drink alcohol or use drugs before you had sexual intercourse the last time you had sex?"
  - Alcohol use
    - "In the past 30 days, on how many days did you have at least one drink of alcohol?"
    - "In the past 30 days, on how many days did you have 5 or more drinks of alcohol in a row, that is, within a couple hours?"


# Discussion, aim 2

- Scale validation
- SPLT task validation
  - Self report relation to SPLT
  - SPLT relation to outcomes
- Self report relation to outcomes
  - Especially as relevant to previous literature
  
------