\school{University of Oregon, Eugene, OR, USA}
\school{University of California, Berkeley, CA, USA}

\degree{Doctor of Philosophy, Psychology, 2018, University of Oregon}
\degree{Master of Science, Psychology, 2014, University of Oregon}
\degree{Bachelor of Arts, Cognitive Science, 2004, University of California, Berkeley}

\interests{Adolescent development}
\interests{Development of individual differences}
\interests{Quantitative methodology}

\position{Graduate Research/Teaching Fellow, University of Oregon, 2012–Present}
\position{Research Coordinator, Stanford University School of Medicine, 2009–2012}
\position{} % extra empty position is neccessary 

\award{Gary E. Smith Summer Professional Development Award, University of Oregon, 2015}
\award{Clarence and Lucille Dunbar Scholarship, University of Oregon, 2014}
\award{} % extra empty award is neccessary 

\publication{Thalmayer, A.G., Saucier, G., Srivastava, S.,  \& Flournoy, J. C. (under review). Ethics-Relevant Values in Adulthood: Longitudinal Findings from the Life and Time Study.}%
\publication{de Macks, Z. A. O., Flannery, J. E., Peake, S. J., Flournoy, J. C., Mobasser, A., Alberti, S. L., ... \& Pfeifer, J. H. (2018). Novel insights from the Yellow Light Game: Safe and risky decisions differentially impact adolescent outcome-related brain function. \textit{NeuroImage}. \href{https://doi.org/10.1016/j.neuroimage.2018.06.058}{10.1016/j.neuroimage.2018.06.058}}%
\publication{Flannery, J. E., Giuliani, N., Flournoy, J. C., \& Pfeifer, J. H. (2017). Neurodevelopmental Changes Across Adolescence in Viewing and Labeling Dynamic Peer Emotions. \textit{Developmental Cognitive Neuroscience, 25}, 113-127. \href{http://dx.doi.org/10.1016/j.dcn.2017.02.003}{10.1016/j.dcn.2017.02.003}}%
\publication{Giuliani, N. R., Flournoy, J. C., Ivie, E. J., Von Hippel, A., \& Pfeifer, J. H. (2017). Presentation and Validation of the DuckEES Child and Adolescent Dynamic Facial Expressions Stimulus Set. \textit{International Journal of Methods in Psychiatric Research, 26(1)}, e1553. \href{https://doi.org/10.1002/mpr.1553}{10.1002/mpr.1553}}%
\publication{King, K. M., Littlefield, A., McCabe, C., Mills, K. L., Flournoy, J. C., \& Chassin, L. (2017). Longitudinal modeling in developmental neuroimaging research: Common challenges, and solutions from developmental psychology. \textit{Developmental Cognitive Neuroscience}. \href{https://doi.org/10.1016/j.dcn.2017.11.009}{10.1016/j.dcn.2017.11.009}}%
\publication{Leonard, J. A., Flournoy, J. C., Lewis-de los Angeles, C. P. \& Whitaker, K. J. (2017). How much motion is too much motion? Determining motion thresholds by sample size for reproducibility in developmental resting-state MRI. \textit{Research Ideas and Outcomes, 3,} e12569. \href{https://doi.org/10.3897/rio.3.e12569}{10.3897/rio.3.e12569}}%
\publication{Madhyastha, T., Peverill, M., Koh, N., McCabe, C., Flournoy, J. C., Mills, K. L., King, K. M., Pfeifer, J. H., \& McLaughlin, K. A. (2017). Current methods and limitations for longitudinal fMRI analysis across development. \textit{Developmental Cognitive Neuroscience}. \href{https://doi.org/10.1016/j.dcn.2017.11.006}{10.1016/j.dcn.2017.11.006}}%
\publication{Matta, T., Flournoy, J. C., \& Byrne, M. (2017). Making an unknown unknown a known unknown: Missing data in longitudinal neuroimaging studies. \textit{Developmental Cognitive Neuroscience}. \href{https://doi.org/10.1016/j.dcn.2017.10.001}{10.1016/j.dcn.2017.10.001}}%
\publication{Flournoy, J. C., Pfeifer, J. H., Moore, W. E., Tackman, A. M., Masten, C. L., Mazziotta, J. C., ... Dapretto, M. (2016). Neural Reactivity to Emotional Faces May Mediate the Relationship Between Childhood Empathy and Adolescent Prosocial Behavior. \textit{Child Development, 87(6),} 1691–1702. \href{https://doi.org/10.1111/cdev.12630}{10.1111/cdev.12630}}%
\publication{Goodkind, M. S., Gallagher-Thompson, D., Thompson, L. W., Kesler, S. R., Anker, L., Flournoy, J., ... O’Hara, R. M. (2015). The impact of executive function on response to cognitive behavioral therapy in late-life depression. \textit{International Journal of Geriatric Psychiatry, 31(4),} 334-339. \href{http://doi.org/10.1002/gps.4325}{10.1002/gps.4325}}%
\publication{O’Hara, R., Marcus, P., Thompson, W. K., Flournoy, J., Vahia, I., Lin, X., ... Jeste, D. V. (2012). 5-HTTLPR Short Allele, Resilience, and Successful Aging in Older Adults. \textit{American Journal of Geriatric Psych, 20(5),} 452-456. \href{http://doi.org/10.1097/JGP.0b013e31823e2d03}{10.1097/JGP.0b013e31823e2d03}}%
\publication{} % extra empty publication is neccessary 
